---
title: "Supplemental Analysis script"
author: "Vincent Somerville"
date: "11/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup ,echo=FALSE}
knitr::opts_chunk$set(message=FALSE, echo=TRUE, eval=FALSE)
.libPaths( c(  .libPaths(),"/home/vincent/R/x86_64-pc-linux-gnu-library/4.0" ) )

.libPaths(c("/home/vincent/R/x86_64-pc-linux-gnu-library/4.1","/home/vincent/R/x86_64-pc-linux-gnu-library/4.0", "/usr/local/lib/R/site-library",                   "/usr/lib/R/site-library"       ,                  "/usr/lib/R/library"                              ))
```

This script includes the complete workflow how the data for the following publication was analysed:

*Extensive diversity and rapid turnover of phage defense systems in cheese-associated bacterial communities*

The script contains the following information. The three parts consist of:

1. Necessary R and bash tools
2. Data collections
3. General script

This is saved and compiled as a Rmarkdown with extensive comments. If questions do arise please contact the lead author. 

# Necessary R libraries 

Here, I load the necessary R libaries

```{r}
library(plyr)
library(tidyverse)
library(ggridges)
library(viridis)
library(ggpmisc)
library(ape)
library(micropan)
library(vegan)
library(ggpubr)
library(rstatix)
library(plotly)
library(ggbiplot)
library(grid)
library(gridExtra)
library(patchwork)
library(ggsignif)
library(stringr)
library(ggplot2)
library(reshape2)
library(ggstance)
library(ggtree)
library(tidytree)
library(treeio)
library(TreeTools)
library(phyloseq)
```

# Necessary bash tools

Here, are the necessary bash tools used throughout the script. 

```{bash}
metaphlan2 --version
fastani --version
blastn --version
prokka --version
cd-hit --version
```


# Necessary data

All data used in the manuscript is deposited on zenodo and NCBI. 

# Genome and metagenome collection

The genomes and metagenomes are deposited on NCBI and can be viewed in the supplmentary table. 

```{r}


genomes_01 <- read_csv("../data_submit/pathway_strain.txt") %>% select(genome,species) %>% unique()
table(genomes_01$species)

genomes_02 <- read_csv("../data_submit/StrainCasType_withSubtype.csv") %>% select(Strain,Organism) %>% unique()
table(genomes_02$Organism)

rbind(genomes_01,genomes_02)
```


# General analysis

Here, we go through the general analysis of the manuscirpt.
The different Results chapter are seperated by the figures. 

# Figure 1

Here, I go through the basic analysis on which the figures are the then created upon (see 02_figure_mainBody_rmk202.Rmd)

## Metagenomic diversity

First we run Metaphlan on the shotgun metagenome dataset. 

```{bash}
###-----------------------------------------------------
####Description section
#Here you can define the locations and file names
###-----------------------------------------------------
threads=10
BaseLocation_output=<directory where your output should go>

###-----------------------------------------------------
####script
###-----------------------------------------------------

rm ${BaseLocation_output}/metaphlan/allStrains_metaphlan.txt

  for bioprojectsss in $(ls /data/projects/p539_crisprscope/0_data/metagenomes) #loop through the different bioprojects
    do
    echo "======================================="
    echo -e "Bioproject: "${bioprojectsss}
    for biosamplesss in $(ls /data/projects/p539_crisprscope/0_data/metagenomes/${bioprojectsss}) #loop through the different biosamples
      do
      echo "======================================="
      echo -e "Biosample: "${biosamplesss}
      
    rm -r ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}
    mkdir -p ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}/{bowtie2ouput,metaphlan_profile}
    
      #check number of read files
      numReadss=$(ls /data/projects/p539_crisprscope/0_data/metagenomes/${bioprojectsss}/${biosamplesss} |wc -l)
    
      if [ "$numReadss" = "2"  ]
        then
        echo -e "there are two read files in the directory, therefore we run a PE metaphlan"
        

   metaphlan2 --nproc 30 --input_type fastq <(zcat /data/projects/p539_crisprscope/0_data/metagenomes/${bioprojectsss}/${biosamplesss}/*_R1*  /data/projects/p539_crisprscope/0_data/metagenomes/${bioprojectsss}/${biosamplesss}/*_R2* ) \
   --bowtie2out  ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}///bowtie2ouput/metagenome_${biosamplesss}_trimmed.bowtie2.bz2 --nproc ${threads} > \
   ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}/metaphlan_profile/profiled_metagenome_${names}_trimmed.txt
   
   grep "s__"  ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}/metaphlan_profile/profiled_metagenome_${names}_trimmed.txt |grep "t__" -v|awk -F "[\t|]" -v culturess="$bioprojectsss" -v strainzzzss="$biosamplesss" '{OFS="\t"}{print culturess,strainzzzss,$7,$8}' >>  ${BaseLocation_output}/metaphlan/allStrains_metaphlan.txt
   
      else
        echo -e "there are only one read file in the directory, therefore we run a single end bwa alignement"
        bwa mem -t 37 $mappingFile \
        /data/projects/p539_crisprscope/0_data/metagenomes/${bioprojectsss}/${biosamplesss}/* | samtools sort -@${threads} -O BAM \
        -o ${BaseLocation_output}/bwaMapping2DB/${bioprojectsss}_${biosamplesss}_rawIllumina_bwamem_sorted.bam - 
        
   metaphlan2 --nproc 30 --input_type fastq <(zcat /data/projects/p539_crisprscope/0_data/metagenomes/${bioprojectsss}/${biosamplesss}/*) \
   --bowtie2out  ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}///bowtie2ouput/metagenome_${biosamplesss}_trimmed.bowtie2.bz2 --nproc ${threads} > \
   ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}/metaphlan_profile/profiled_metagenome_${names}_trimmed.txt
   
   grep "s__"  ${BaseLocation_output}/metaphlan/${bioprojectsss}_${biosamplesss}/metaphlan_profile/profiled_metagenome_${names}_trimmed.txt |grep "t__" -v|awk -F "[\t|]" -v culturess="$bioprojectsss" -v strainzzzss="$biosamplesss" '{OFS="\t"}{print culturess,strainzzzss,$7,$8}' >>  ${BaseLocation_output}/metaphlan/allStrains_metaphlan.txt
   
        
      fi
  
    done #biosamplessss
  done #bioprojectsss

```

This is the analysis of the sequenced metagenomic datasets. 

```{r}


myFMBN_new1_prev_ab_table <- read_delim("data_submit/20210426_final_all1_prev_ab_table.txt", "\t", escape_double = FALSE, trim_ws = TRUE) %>% filter(!is.na(species))

myFMBN_new_again_again2_filt_OTU_table <- read_delim("data_submit/20210426_final_all1_filt_OTU_table.txt","\t", escape_double = FALSE, trim_ws = TRUE)


names.use <- names(myFMBN_new_again_again2_filt_OTU_table)[(names(myFMBN_new_again_again2_filt_OTU_table) %in% c("sample_name",myFMBN_new1_prev_ab_table$label))]
OTU_sample <- (myFMBN_new_again_again2_filt_OTU_table)[, names.use] 

OTU_sample_long <- gather(OTU_sample, species,relAbundance,colnames(OTU_sample)[2]:colnames(OTU_sample)[ncol(OTU_sample)],  factor_key=TRUE,na.rm = TRUE) %>% filter(relAbundance>0.1)
OTU_sample_long$sample_name %>% unique() %>% length()
OTU_sample_long_summarise <- OTU_sample_long%>%  group_by(species)  %>%  dplyr::summarize(prevelance=100*(n()/length(unique(OTU_sample_long$sample_name))),median = 100*(median(relAbundance))) 

OTU_sample_long_summarise %>% filter(prevelance>2&median<0.1)
OTU_sample_long_summarise %>% filter(prevelance>2&median>0.1)

final_species <- OTU_sample_long_summarise %>% filter(prevelance>2&median>0.1)
##----------------------------------------------
###SUBSET PROPERLY

dat <- readRDS("data_submit/20210426_final_all1.RDS")
studies_DF <- dat$studies

tmp <- studies_DF %>% select(studyId,short_descr)

samples_DF <- dat$samples

##-----------------
#select only specific cheese samples and feremented spoilage
##-----------------

samples_selected <- samples_DF %>% filter(foodId %in% c("A02QJ",
"A02ZQ",
"A02ZS",
"A048Y",
"A02YH",
"A02TY",
"A02QL",
"A02XL",
"A02ZL",
"A02ZL",
"A02SV",
"A02XE",
"A02YM",
"A02RH",
"A02ST",
"A02YE",
"A02YD",
"A02YM",
"A02QP",
"A02TJ",
"A02XH",
"A030V",
"A02RH",
"A02RN",
"A02RL",
"A02RJ",
"A030Y",
"A02SF",
"A030X",
"A02RR",
"A02TY",
"A02VA",
"A02SV",
"A04NV",
"A02RV",
"A02YK",
"A02YP",
"A02VF",
"A02ZD",
"A02YE",
"A02ZN",
"A02XE",
"A02YE")) %>% filter(spoilage=="Fermented") %>% dplyr::filter(!grepl("rind",description, ignore.case =TRUE))

studyextension <- samples_selected %>% select(studyId,label_2)
names_selected <- samples_selected$label_2
OTU_subset <- OTU_sample %>% filter(sample_name %in%names_selected)

##-----------------------------------
#calculate abundance
##-----------------------------------
OTU_subset_long <- gather(OTU_subset, species,relAbundance,colnames(OTU_subset)[2]:colnames(OTU_subset)[ncol(OTU_subset)],  factor_key=TRUE,na.rm = TRUE) %>% filter(relAbundance>0.2) #%>% filter(relAbundance!=0)
OTU_subset_long_summarise <- OTU_subset_long%>%  group_by(species)  %>%  dplyr::summarize(prevelance=100*(n()/length(unique(OTU_subset_long$sample_name))),median = 100*(median(relAbundance))) 
final_species_final <- OTU_subset_long_summarise %>% filter(prevelance>2&median>0.1)

##-----------------------------------
#prepare shotgun metaG data
##-----------------------------------
X20210424_metaphlan <- read_csv("data_submit/20210424_metaphlan.txt") %>% select(-X1)%>% rename("relAbundance"="abundance") %>% rename("sample_name"="SRA_ID")  %>% rename("studyId"="ProjectID") %>% add_column(method="metagenome") #%>% filter(relAbundance>0.2) 
OTU_subset_long <- gather(OTU_subset, species,relAbundance,colnames(OTU_subset)[2]:colnames(OTU_subset)[ncol(OTU_subset)],  factor_key=TRUE,na.rm = TRUE)  %>% mutate(relAbundance=100*relAbundance) #%>% filter(relAbundance>0.2)#%>% filter(relAbundance!=0)
OTU_subset_long_extended <- merge(OTU_subset_long,studyextension,by.x = "sample_name",by.y="label_2") %>% as.tibble()  %>% select(studyId,sample_name,species,relAbundance)%>% add_column(method="16S")
OTU_subset_long_extended_full <- rbind(X20210424_metaphlan,OTU_subset_long_extended) %>% filter(species!=0)

ggplot(OTU_subset_long_extended_full, aes(x=sample_name,y=relAbundance,color=species,fill=species))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")

##------color subdominant

OTU_subset_long_extended_full_subdominant <- OTU_subset_long_extended_full %>% filter(relAbundance<0.2) %>% mutate(species_new="subdominant")
OTU_subset_long_extended_full_dominant <- OTU_subset_long_extended_full %>% filter(relAbundance>=0.2) %>% mutate(species_new=species)
OTU_subset_long_extended_full_dominance <- rbind(OTU_subset_long_extended_full_dominant,OTU_subset_long_extended_full_subdominant)  %>% filter(relAbundance!=0)%>% droplevels()

ggplot(OTU_subset_long_extended_full_dominance, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")

##------remove samples with less then 50% species identification


tmp_samples <- OTU_subset_long_extended_full_dominance %>% group_by(sample_name) %>% summarise(sumsss =sum(relAbundance)) %>% filter(sumsss<50)
OTU_subset_long_extended_full_dominance_filtered <- OTU_subset_long_extended_full_dominance %>% filter(!sample_name %in% tmp_samples$sample_name)%>% droplevels()
ggplot(OTU_subset_long_extended_full_dominance_filtered, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")

##------summarise to remove low prevelence species

OTU_subset_long_summarise <- OTU_subset_long_extended_full_dominance_filtered%>%  group_by(species)  %>%  dplyr::summarize(prevelance=100*(n()/length(unique(OTU_subset_long_extended_full_dominance_filtered$sample_name))),median =(median(relAbundance))) 
OTU_subset_long_summarise_prevelance <- OTU_subset_long_summarise %>% filter(prevelance>0.5)%>% filter(median>0.25)
OTU_subset_long_summarise__subdominant <- OTU_subset_long_summarise %>% filter(!prevelance>0.5)%>% filter(!median>0.25)


OTU_subset_long_extended_full_dominance_dom <- OTU_subset_long_extended_full_dominance_filtered %>% filter(species_new %in% OTU_subset_long_summarise_prevelance$species)
OTU_subset_long_extended_full_dominance_sub <- OTU_subset_long_extended_full_dominance_filtered %>% filter(!species_new %in% OTU_subset_long_summarise_prevelance$species)
OTU_subset_long_extended_full_dominance_sub$species_new <- "subdominant"

OTU_subset_long_extended_full_dominance_again <- rbind(OTU_subset_long_extended_full_dominance_dom,OTU_subset_long_extended_full_dominance_sub)  %>% filter(relAbundance!=0)%>% droplevels() %>% filter(species_new!="Pediococcus sp")
ggplot(OTU_subset_long_extended_full_dominance_again, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")

###add to 100% for unassigned
unassigned <- OTU_subset_long_extended_full_dominance_again %>% group_by(studyId,sample_name,method) %>% summarise(relAbundance=100-sum(relAbundance)) %>% filter(relAbundance>0.1) %>% mutate(species="unassigned")%>% mutate(species_new="unassigned") %>% select(studyId,sample_name,species,relAbundance,method,species_new)

OTU_subset_long_extended_full_dominance_again_final <- rbind(OTU_subset_long_extended_full_dominance_again,unassigned)  %>% droplevels() 
ggplot(OTU_subset_long_extended_full_dominance_again_final, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")

###------------------------------------------
##order species

ordersss <- OTU_subset_long_summarise_prevelance[order(OTU_subset_long_summarise_prevelance$prevelance),] %>% select(species) %>% pull(species)
ordersss_final <- c("unassigned","subdominant",ordersss)
OTU_subset_long_extended_full_dominance_again_final$species_new = factor(OTU_subset_long_extended_full_dominance_again_final$species_new, levels=ordersss_final)
ggplot(OTU_subset_long_extended_full_dominance_again_final, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")

colorss <- c(
  "grey77", #unassigned
  "grey66", #subdominant
  rev(rainbow(23)[-1]),
  #...
  "#87deaa", #llactis
  "#ffcc00", #lhelv
  "#fdc2b5", #ldel
  "#007fa6") #Sterm

alphasss <- c(rep(0.3,24),rep(1,4))

ggplot(OTU_subset_long_extended_full_dominance_again_final, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new,alpha=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")+
  scale_color_manual(values = colorss)+ scale_fill_manual(values = colorss)+scale_alpha_manual(values = alphasss)

OTUs_print <- OTU_subset_long_extended_full_dominance_again_final %>% group_by(species_new)    %>%  dplyr::summarize(counts=n(),prevelance=100*(n()/length(unique(OTU_subset_long_extended_full_dominance_again_final$sample_name))),median = median(relAbundance)) %>% filter(!species_new %in% c("unassigned","subdominant"))

# write.csv(OTUs_print,file="~/Desktop/Projects/2020_CRISPRscope/metaphlan/20210428_metaphlan_final_species.txt")

##====================================
###color accordingly..
##====================================

data.frame(species= c( "Lactobacillus acidophilus",
"Lactobacillus fermentum", "Pediococcus parvulus",
"Bifidobacterium animalis", "Brevibacterium aurantiacum", "Lactobacillus versmoldensis",
"Leuconostoc lactis", "Corynebacterium casei", "Pediococcus pentosaceus",
"Propionibacterium freudenreichii", "Streptococcus luteciae",
"Lactobacillus rhamnosus", "Lactobacillus coryniformis", "Lactobacillus curvatus",
"Brevibacterium linens", "Leuconostoc pseudomesenteroides", "Lactobacillus plantarum",
"Staphylococcus sciuri", "Leuconostoc mesenteroides", "Propionibacterium acnes",
"Staphylococcus equorum", "Lactobacillus casei", "Lactococcus lactis",
"Lactobacillus helveticus", "Lactobacillus delbrueckii", "Streptococcus thermophilus"
),
color= c("#fdc2b496","#a4778bff","#a3758a64","#17c3b2ff","#16c3b396","#a7c5da4b","#ff6d584b","grey","#a3768aaf","#61988eff","#007ea796","#a7c5da96","#eeabc4ff","#eeaac396","#14c2b14b","#ff6f59ff","#a3758a32","#c8beb7ff","#ff6e5896","grey","grey","#a7bed3ff","#eeba0bff","#fcc2b44b","#fdc2b5ff","#007fa6ff"))

speciesss <- c("unassigned", "subdominant", "Lactobacillus acidophilus", 
"Lactobacillus fermentum", "Pediococcus parvulus", "Pediococcus sp", 
"Bifidobacterium animalis", "Brevibacterium aurantiacum", "Lactobacillus versmoldensis", 
"Leuconostoc lactis", "Corynebacterium casei", "Pediococcus pentosaceus", 
"Propionibacterium freudenreichii", "Streptococcus luteciae", 
"Lactobacillus rhamnosus", "Lactobacillus coryniformis", "Lactobacillus curvatus", 
"Brevibacterium linens", "Leuconostoc pseudomesenteroides", "Lactobacillus plantarum", 
"Staphylococcus sciuri", "Leuconostoc mesenteroides", "Propionibacterium acnes", 
"Staphylococcus equorum", "Lactobacillus zeae", "Lactococcus lactis", 
"Lactobacillus helveticus", "Lactobacillus delbrueckii", "Streptococcus thermophilus"
)

colorsss <- c("NA","NA","#fdc2b496","#a4778bff","#a3758a64","#17c3b2ff","#16c3b396","#a7c5da4b","#ff6d584b","grey","#a3768aaf","#61988eff","#007ea796","#a7c5da96","#eeabc4ff","#eeaac396","#14c2b14b","#ff6f59ff","#a3758a32","#c8beb7ff","#ff6e5896","grey","grey","#a7bed3ff","#eeba0bff","#fcc2b44b","#fdc2b5ff","#007fa6ff")

##------------filter out samples with only two samples
tmp <- OTU_subset_long_extended_full_dominance_again_final %>% select(studyId,sample_name) %>% as.data.frame()%>% unique()
OTU_subset_long_extended_full_dominance_again_final <- OTU_subset_long_extended_full_dominance_again_final %>% filter(!studyId %in% c("PRJNA286900","ST107","ST23","ST80"))
##------------plot

ggplot(OTU_subset_long_extended_full_dominance_again_final, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")+
  scale_color_manual(values = colorsss)+ scale_fill_manual(values = colorsss)

##------------------------------order according to mesophilic/thermophilus
OTU_subset_long_extended_full_dominance_again_final$studyId = factor(OTU_subset_long_extended_full_dominance_again_final$studyId, levels=c("20201217_metagenomes", "PRJNA603575", "ST3", "ST4", "ST7", "ST10", 
"ST9", "ST18", "ST22", "ST8", "ST48", "ST49",
 "CheeseRaclette", "PRJEB23938", "PRJEB30079", 
"PRJEB32768","ST106","ST6"))

##------------plot

ggplot(OTU_subset_long_extended_full_dominance_again_final, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")+
  scale_color_manual(values = colorsss)+ scale_fill_manual(values = colorsss)+theme(legend.position = "none",axis.ticks.x = element_blank(),axis.text.x=element_blank(),strip.background = element_blank(),strip.text.x = element_text(angle = 45, 
    size = 8))+ scale_y_continuous(expand = c(0,0))+labs(x="samples",y="relative Abundance [%]")

##-------------refshort

studies_DF_shrot <- studies_DF %>% select(c("studyId","ref_short"))
namechange <- setNames(studies_DF_shrot$ref_short,studies_DF_shrot$studyId)

OTU_subset_long_extended_full_dominance_again_final$studyId <- plyr::revalue(OTU_subset_long_extended_full_dominance_again_final$studyId,namechange)
OTU_subset_long_extended_full_dominance_again_final$studyId <- plyr::revalue(OTU_subset_long_extended_full_dominance_again_final$studyId,c("20201217_metagenomes"="Swiss starter cultures","CheeseRaclette"="Swiss Raclette"))
OTU_subset_long_extended_full_dominance_again_final$studyId <- plyr::revalue(OTU_subset_long_extended_full_dominance_again_final$studyId,c("PRJNA603575"="Pasolli et al., 2020","PRJEB23938"="Duru et al., 2018","PRJEB30079"="Lordan et al., 2019","PRJEB32768"="Walsh et al., 2020"))

##------------plot
levels(OTU_subset_long_extended_full_dominance_again_final$species_new) %>% length()
colorsss %>% length()

final_abundanc_barchart <- ggplot(OTU_subset_long_extended_full_dominance_again_final, aes(x=sample_name,y=relAbundance,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")+
  scale_color_manual(values = colorsss)+ scale_fill_manual(values = colorsss)+theme(legend.position = "none",axis.ticks.x = element_blank(),axis.text.x=element_blank(),strip.background = element_blank(),strip.text.x = element_text(angle = 45, 
    size = 8))+ scale_y_continuous(expand = c(0,0))+labs(x="samples",y="relative Abundance [%]")+ coord_cartesian(clip = 'off')

   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/species_barplot_abundance.pdf",width=22,height=6)

final_abundanc_barchart
# dev.off()

##-----------------------------------------------------------
##------------make 100% relative abundance
##-----------------------------------------------------------
OTU_subset_long_extended_full_dominance_again_final_tmp <- OTU_subset_long_extended_full_dominance_again_final %>% filter(!species_new %in% c("unassigned","subdominant"))
tmp1 <- OTU_subset_long_extended_full_dominance_again_final_tmp %>% group_by(sample_name) %>% summarise(total_abundance=sum(relAbundance))
OTU_subset_long_extended_full_dominance_again_final_02 <- merge(OTU_subset_long_extended_full_dominance_again_final_tmp,tmp1,by="sample_name") %>% mutate(relAbundance_02=100*(relAbundance/total_abundance))

colorzzz <-  c("#fdc2b496","#a4778bff","#a3758a64","#17c3b2ff","#16c3b396","#a7c5da4b","#ff6d584b","grey","#a3768aaf","#61988eff","#007ea796","#a7c5da96","#eeabc4ff","#eeaac396","#14c2b14b","#ff6f59ff","#a3758a32","#c8beb7ff","#ff6e5896","grey","grey","#a7bed3ff","#eeba0bff","#fcc2b44b","#fdc2b5ff","#007fa6ff")

final_abundanc_barchart_02 <- ggplot(OTU_subset_long_extended_full_dominance_again_final_02, aes(x=sample_name,y=relAbundance_02,color=species_new,fill=species_new))+geom_bar(stat = "identity")+theme_classic()+facet_grid(~studyId, scales = "free_x", space = "free_x")+
  scale_color_manual(values = colorzzz)+ scale_fill_manual(values = colorzzz)+theme(legend.position = "none",axis.ticks.x = element_blank(),axis.text.x=element_blank(),strip.background = element_blank(),strip.text.x = element_text(angle = 45, 
    size = 8))+ scale_y_continuous(expand = c(0,0))+labs(x="samples",y="relative Abundance [%]")+ coord_cartesian(clip = 'off')
final_abundanc_barchart_02
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/species_barplot_abundance.pdf",width=22,height=6)

final_abundanc_barchart_02
# dev.off()


# write.table(OTU_subset_long_extended_full_dominance_again_final_02,"~/Desktop/Projects/2020_CRISPRscope/metagenomes_sample_count.txt",sep=",",quote = FALSE,row.names = FALSE)


##-----------------------------------------------------------
##------------create color dataframe
##-----------------------------------------------------------

namesss <- levels(OTU_subset_long_extended_full_dominance_again_final$species_new) 
namesss <- namesss[namesss!="Pediococcus sp"]

color_df <- data.frame(species=namesss,color=colorsss)
# write.table(color_df,"~/Desktop/Projects/2020_CRISPRscope/species_colors.txt",sep="\t",quote = FALSE,row.names = FALSE)

```

## Species phylogeny

First step includes the annotation of the genomes.

```{bash}
conda activate PROKKA
  num=1

 for genomessss in $(ls /data/projects/p539_crisprscope/06_genomes_speciesOFInterest// |grep ".fna$" |sed 's/.fna//g')
 do
 echo $genomessss
 
 rm -r /data/projects/p539_crisprscope/02_PROKKA_speciesOFInterest/${genomessss}
 prokka --cpus  35 --outdir /data/projects/p539_crisprscope/02_PROKKA_speciesOFInterest/${genomessss} --locustag REFs_${num}  --proteins proteins.faa --evalue 0.001 --addgenes /data/projects/p539_crisprscope/06_genomes_speciesOFInterest/$genomessss.fna
 num=$((num+1))

  done
```

Then we create a phylogeny for the different species with BCGtree.

```{bash}
###------------------------------------
#outgroup
###------------------------------------
mkdir -p /data/projects/p539_crisprscope/01_data/phylogeny_outgroup

###CBASS
grep "eggerthia catenaformis" -i  /data/projects/p539_crisprscope/ncbi/genome/download.file | tail -1 | awk 'BEGIN{FS="\t";OFS="\t"}  {print $20}' |sed -r 's|(ftp://ftp.ncbi.nlm.nih.gov/genomes/all/.+/)(GCF_.+)|\1\2\/\2_protein.faa.gz|' >  /data/projects/p539_crisprscope/01_data/phylogeny_outgroup/eggerthia_catenaformis.log
cd /data/projects/p539_crisprscope/01_data/phylogeny_outgroup/
wget -i /data/projects/p539_crisprscope/01_data/phylogeny_outgroup/eggerthia_catenaformis.log
gunzip *.faa.gz
mv GCF_002148845.1_ASM214884v1_protein.faa eggerthia_catenaformis.faa

##----------------------------------------
##prep file
##----------------------------------------
rm -r /data/projects/p539_crisprscope/05_BCGtree_new
mkdir -p /data/projects/p539_crisprscope/05_BCGtree_new
rm /data/projects/p539_crisprscope/05_BCGtree_new/log.txt

echo -e "--proteome Outgroup=/data/projects/p539_crisprscope/01_data/phylogeny_outgroup/eggerthia_catenaformis.faa" >> /data/projects/p539_crisprscope/05_BCGtree_new/log.txt

for speciesss in $(ls /data/projects/p539_crisprscope/02_PROKKA_speciesOFInterest/ )
do
echo "===================================="
echo ${speciesss}

prokka_tmp=$(ls /data/projects/p539_crisprscope/02_PROKKA_speciesOFInterest/${speciesss}/ |grep ".faa" )

echo -e "--proteome "${speciesss}"=/data/projects/p539_crisprscope/02_PROKKA_speciesOFInterest/"${speciesss}"/"${prokka_tmp} >> /data/projects/p539_crisprscope/05_BCGtree_new/log.txt
head -1 /data/projects/p539_crisprscope/02_PROKKA_speciesOFInterest/${speciesss}//${prokka_tmp}
done

echo -e "--outdir /data/projects/p539_crisprscope/05_BCGtree_new/" >> /data/projects/p539_crisprscope/05_BCGtree_new/log.txt
wc -l /data/projects/p539_crisprscope/05_BCGtree_new/log.txt
##----------------------------------------
##run bcgTree
##----------------------------------------
bcgTree.pl @/data/projects/p539_crisprscope/05_BCGtree_new/log.txt --threads 37
```

make nice tree in R.

```{r }


Braod_species_tree_raw <- read.nexus("data_submit/RAxML_bestTree_cleaned.final", tree.names = NULL, force.multi = FALSE)
Braod_species_tree_raw$tip.label
##--------START: changing names
Renaming_lactobacillus_genus <- read_csv("data_submit/Renaming_lactobacillus_genus.csv")
namechange <- setNames(Renaming_lactobacillus_genus$new_name,Renaming_lactobacillus_genus$old_name_with)

Braod_species_tree_raw$tip.label <- plyr::revalue(Braod_species_tree_raw$tip.label,namechange)
Braod_species_tree_raw$tip.label <- gsub("_", " " ,Braod_species_tree_raw$tip.label)
##--------END: changing names


gtree_01 <- ggtree(Braod_species_tree_raw) +
  geom_tiplab(align = TRUE)+ coord_cartesian(clip = 'off') + #+ xlim(0, 1.8)+ # tiplabls aligned
  geom_rootedge(rootedge = 0) +geom_treescale() + theme_tree2() 
 #NO ROOT  
  
gtree_01
```

### Additional species info

```{r }
##------------------------------------
#Abundance 
##------------------------------------

X20210428_metaphlan_final_species <- read_csv("data_submit/20210428_metaphlan_final_species.txt") %>% select(-counts)%>% select(-X1)  %>% rename("id"="species_new") 
X20210428_metaphlan_final_species <- rbind(X20210428_metaphlan_final_species,c("Outgroup",0,0)) %>% as.data.frame()

##--------START: changing names
namechange <- setNames(Renaming_lactobacillus_genus$new_name,Renaming_lactobacillus_genus$old_name)

X20210428_metaphlan_final_species$id <- plyr::revalue(X20210428_metaphlan_final_species$id,namechange)
X20210428_metaphlan_final_species$id <- gsub("_", " " ,X20210428_metaphlan_final_species$id)
X20210428_metaphlan_final_species$id <- plyr::revalue(X20210428_metaphlan_final_species$id,c("Streptococcus luteciae"="Streptococcus lutetiensis","Staphylococcus sciuri"="Mammaliicoccus sciuri","Propionibacterium acnes"="Cutibacterium acnes"))

##--------END: changing names
str(X20210428_metaphlan_final_species)
Braod_species_tree_raw$tip.label %in% X20210428_metaphlan_final_species$id
X20210428_metaphlan_final_species[which(!X20210428_metaphlan_final_species$id %in% Braod_species_tree_raw$tip.label),]
Braod_species_tree_raw$tip.label[which(!Braod_species_tree_raw$tip.label %in% X20210428_metaphlan_final_species$id)]



X20210428_metaphlan_final_species$median <- as.numeric(X20210428_metaphlan_final_species$median)

p4 <- facet_plot(gtree_01, panel = 'Abundance', data = X20210428_metaphlan_final_species, 
                geom = geom_barh, 
                mapping = aes(x = median), width = 0.8, 
                stat='identity' ) +xlim_expand(c(0,100), "Abundance")
p4

##------------------------------------
#Prevalance 
##------------------------------------

X20210428_metaphlan_final_species$prevelance <- as.numeric(X20210428_metaphlan_final_species$prevelance)

p5 <- facet_plot(p4, panel = 'Prevelance', data = X20210428_metaphlan_final_species, 
                geom = geom_barh, 
                mapping = aes(x = prevelance), width = 0.8, 
                stat='identity' )+xlim_expand(c(0,100), "Prevelance") #+ xlim(0, 100)
p5
##-------------------------------------------
#genome count
##-------------------------------------------

species_names <- read_delim("data_submit/genome_count_SEL26.txt",  ";", escape_double = FALSE, col_names = c("value", "id"),  trim_ws = TRUE, skip = 1) %>% select(id)

genome_count_SEL186 <- read_csv("data_submit/export_genome_count_SEL185.csv", col_names =c("X1","value", "id"),  trim_ws = TRUE, skip = 1) %>% select(-X1) %>% add_column(category="NCBI")
export_genome_count_SELDIALACT <- read_csv("data_submit/export_genome_count_SELDIALACT.csv", col_names = c("X1","value", "id"), skip = 1)  %>% select(-X1) %>% add_column(category="Dialact")

genome_count_SEL186_final <- rbind(genome_count_SEL186,export_genome_count_SELDIALACT) %>% as.data.frame() %>% select(id,value,category)

genome_count_SEL26_final <- genome_count_SEL186_final %>% filter(id %in% c("Leuconostoc_lactis","Streptococcus_thermophilus", "Lactobacillus_helveticus", "Lactococcus_lactis", "Brevibacterium_aurantiacum","Lactobacillus_acidophilus","Cutibacterium_acnes","Mammaliicoccus_sciuri" ,"Lactobacillus_delbrueckii" ,       "Leuconostoc_mesenteroides","Propionibacterium_freudenreichii" ,"Pediococcus_pentosaceus","Staphylococcus_equorum","Brevibacterium_linens", "Streptococcus_lutetiensis", "Corynebacterium_casei" ,"Bifidobacterium_animalis",  "Pediococcus_parvulus" ,"Leuconostoc_pseudomesenteroides", "Loigolactobacillus_coryniformis","Lacticaseibacillus_rhamnosus","Lacticaseibacillus_casei","Lactiplantibacillus_plantarum" ,"Limosilactobacillus_fermentum","Companilactobacillus_versmoldensis","Latilactobacillus_curvatus")
)

##--------START: changing names
namechange <- setNames(Renaming_lactobacillus_genus$new_name,Renaming_lactobacillus_genus$old_name_with)

genome_count_SEL26_final$id <- plyr::revalue(genome_count_SEL26_final$id,namechange)
genome_count_SEL26_final$id <- gsub("_", " " ,genome_count_SEL26_final$id)

##--------END: changing names

Braod_species_tree_raw$tip.label %in% genome_count_SEL26_final$id
genome_count_SEL26_final$id %in% Braod_species_tree_raw$tip.label

genome_count_SEL26_final$value <- as.numeric(genome_count_SEL26_final$value)

p6 <- facet_plot(p5, panel = 'Genome count', data = genome_count_SEL26_final, 
                geom = geom_barh, 
                mapping = aes(x = value,fill=category), width = 0.8, 
                stat='identity' ) #+xlim_expand(c(0,500), 'Genome count')
p6

   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/species_tree_abundance_genome_count.pdf",width=10,height=6)

# p6
# dev.off()

```

## FastaANI

Here, I calculate the Average nucleotide identity (ANI) between any two given genomes of all analysed species 

```{bash}

###---------------------------
##make a list of locations of all Sterm and ldel genomes to comapre
###---------------------------

##-------------------------------all


rm -r /data/projects/p539_crisprscope/fastani/
mkdir -p /data/projects/p539_crisprscope/fastani/{20210123_run,log}
for DBss in $(ls /data/projects/p539_crisprscope/01_data/20210123_assemblies/)
do

for datesss in $(ls /data/projects/p539_crisprscope/01_data/20210123_assemblies/${DBss}/)
do

#rm  /data/projects/p539_crisprscope/fastani/log/filenames_all_fastani.txt
for speciesss in $(ls /data/projects/p539_crisprscope/01_data/20210123_assemblies/${DBss}/${datesss})
do
echo "===================================="
echo ${speciesss}
 ls /data/projects/p539_crisprscope/01_data/20210123_assemblies/${DBss}/${datesss}/${speciesss} |wc -l
 ls /data/projects/p539_crisprscope/01_data/20210123_assemblies/${DBss}/${datesss}/${speciesss} |grep -c ".fna"
 for genomessss in $(ls /data/projects/p539_crisprscope/01_data/20210123_assemblies/${DBss}/${datesss}/${speciesss})
 do
 
 genome_short=$(echo ${genomessss}|sed 's/.fna//g')
echo -e "/data/projects/p539_crisprscope/01_data/20210123_assemblies/"${DBss}"/"${datesss}"/"${speciesss}"/"${genome_short}".fna"  >> /data/projects/p539_crisprscope/fastani/log/filenames_all_fastani.txt
echo -e ${speciesss}"\t"${genome_short}  >> /data/projects/p539_crisprscope/fastani/log/filenames_all_description_genomes.txt

  done
done
done
done

wc -l /data/projects/p539_crisprscope/fastani/log/filenames_all_description_genomes.txt
wc -l /data/projects/p539_crisprscope/fastani/log/filenames_all_fastani.txt


###---------------------------
##fastaANI
###---------------------------
fastANI --fragLen 1000 -t 37 --ql /data/projects/p539_crisprscope/fastani/log/filenames_all_fastani.txt \
  --rl /data/projects/p539_crisprscope/fastani/log/filenames_all_fastani.txt \
  -o /data/projects/p539_crisprscope/fastani/20210123_run/20201219_ANI_output_all.txt

```

```{r}

##----------------
##sterm
##----------------

fastani_thibualt <- read_delim("data_submit/20201219_ANI_output_all.txt", "\t", escape_double = FALSE, col_names = FALSE,  trim_ws = TRUE) 
genomeDescription <- read_delim("data_submit/filenames_all_description_genomes.txt",  "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)

table(genomeDescription$X1)
table(genomeDescription$X1)
genomeDescription$X1 <- plyr::revalue(genomeDescription$X1,c("Lactobacillus_delbrueckii_subsp_bulgaricus"="Lactobacillus_delbrueckii","Lactobacillus_delbrueckii_subsp_lactis"="Lactobacillus_delbrueckii","Lactococcus_lactis_subsp_lactis"="Lactococcus_lactis","Lactococcus_lactis_subsp_cremoris"="Lactococcus_lactis"))
fastani_thibualt$AF <- (100*(fastani_thibualt$X4/fastani_thibualt$X5))


fastani_thibualt$genome1 <- str_split_fixed(fastani_thibualt$X1, fixed("/"), 9)[,8] %>%  gsub(".fna","",.)
fastani_thibualt$genome2 <- str_split_fixed(fastani_thibualt$X2, fixed("/"), 9)[,8] %>%  gsub(".fna","",.)

fastani_thibualt_cleaned <- fastani_thibualt %>% rename("X3"="id") %>% select(c(genome1,genome2,id,AF))
tmp1 <- merge(fastani_thibualt_cleaned,genomeDescription,by.x="genome1",by.y="X2") %>% rename("X1"="species_genome1")
fastani_thibualt_final <- merge(tmp1,genomeDescription,by.x="genome2",by.y="X2") %>% rename("X1"="species_genome2")


 fastani_thibualt_final%>%  group_by(species_genome1,species_genome2)  %>%  dplyr::summarize(median = median(id)) 



```

```{bash}
count=0
for speciesss in $(ls /data/projects/p539_crisprscope/01_data/20201219_all)
do
echo "===================================="
echo ${speciesss}
ls /data/projects/p539_crisprscope/01_data/20201219_all/${speciesss}/ |wc -l
zss=$(ls /data/projects/p539_crisprscope/01_data/20201219_all/${speciesss}/ |wc -l)

 count=$((count+zss))

done
echo $count
```

## Defense mechanism diversity

In the bioinfics scirpt you find how we identified the defense mechanisms. 
Here, we show the analysis of the data. 

```{r }
pathway_identification_species <- read_csv("data_submit/pathway_species.txt")

#order the pathways according to these
pathway_identification_species$mechanism_final  = factor(pathway_identification_species$mechanism_final , levels=c(
  "R-M-I", "R-M-II", "R-M-IIG", "R-M-III", "R-M-IV",
  "BREX","Disarm-I","Disarm-II", 
  "Druantia-I",  "Druantia-II","Druantia-III","Hachiman","Kiwa","Lamassu","Shedu","Thoeris","Zorya-I","Zorya-II","Wadjet-I","Septu",
  "CBASS","Abi",
  "GTPase","dNTPaminase","Retrons"
  
))

##-------------------------------

ggheatmap <- ggplot(pathway_identification_species, aes(y=species, x=mechanism_final, fill = percent_species))+
 geom_tile(color = "white",size=1.1)+
 scale_fill_gradient2(low = "white", high = "red", mid = "white",
   midpoint = 0, limit = c(-100,100), space = "Lab",
   name="max growth") +theme_classic()+
  #theme_minimal()+ # minimal theme
  scale_x_discrete(position = "top") +
  # labs(x=expression(italic("S. thermophilus")),y=expression(italic("L. delbrueckii")))+
 theme(axis.text.y = element_text(size=7),
      axis.text.x = element_text(angle = 45,vjust=1,
    size = 12, hjust = 0),axis.title=element_blank(), 
    legend.title = element_blank())
 # coord_fixed()
# Print the heatmap
 print(ggheatmap)
  
##=========================
##add CRISPR
##=========================
pathway_identification_species_plot <- pathway_identification_species %>% select(species,mechanism_final,percent_species)

StrainCasType_subtypes <- read_csv("data_submit/StrainCasType_withSubtype.csv") %>% dplyr::rename(., "species"="Organism")%>% mutate(mechanism_final=paste0("CRISPR type ",Type,"_",Subtype)) %>% select(species,Strain,mechanism_final)%>% filter(mechanism_final!="CRISPR type III_B")

StrainCasType_general_CRISPR <- read_csv("data_submit/StrainCasType_withSubtype.csv") %>% rename(., "species"="Organism")%>% mutate(mechanism_final="CRISPR") %>% select(species,Strain,mechanism_final)

StrainCasType <- rbind(StrainCasType_subtypes,StrainCasType_general_CRISPR)

StrainCasType <- StrainCasType %>% group_by(species,Strain,mechanism_final) %>% summarise(count=n() ) %>% select(-count)
StrainCasType_counts <- StrainCasType %>% group_by(species,mechanism_final) %>% summarise(counts=n())

StrainCasType_final <- merge(StrainCasType_counts,species_counts,by="species") %>% mutate(percent_species=100*(counts.x/counts.y))%>% select(species,mechanism_final,percent_species)

Strain_all_Type_final <- rbind(pathway_identification_species_plot,StrainCasType_final)
Strain_all_Type_final$mechanism_final  = factor(Strain_all_Type_final$mechanism_final , levels=c(
  "R-M-I", "R-M-II", "R-M-IIG", "R-M-III", "R-M-IV",
  "BREX","Disarm-I","Disarm-II", 
  "Druantia-I",  "Druantia-II","Druantia-III","Hachiman","Kiwa","Lamassu","Shedu","Thoeris","Zorya-I","Zorya-II","Wadjet-I","Septu",
  "CBASS","Abi",
  "GTPase","dNTPaminase","Retrons",
  #"CRISPR type I",  "CRISPR type II", "CRISPR type III"
  "CRISPR","CRISPR type I_B","CRISPR type I_C","CRISPR type I_E","CRISPR type I_G","CRISPR type II_A","CRISPR type II_C","CRISPR type III_A"
))

##--------START: changing names
Renaming_lactobacillus_genus <- read_csv("data_submit/Renaming_lactobacillus_genus.csv")
namechange <- setNames(Renaming_lactobacillus_genus$new_name,Renaming_lactobacillus_genus$old_name_with)

Strain_all_Type_final$species <- plyr::revalue(Strain_all_Type_final$species,namechange)
Strain_all_Type_final$species <- gsub("_", " " ,Strain_all_Type_final$species)

##--------END: changing names

Strain_all_Type_final %>% filter(mechanism_final=="CRISPR")

ggheatmap <- ggplot(Strain_all_Type_final, aes(y=species, x=mechanism_final, fill = percent_species))+
 geom_tile(color = "white",size=1.1)+
 scale_fill_gradient2(low = "white", high = "red", mid = "white",
   midpoint = 0, limit = c(-100,100), space = "Lab",
   name="max growth") +theme_classic()+
  #theme_minimal()+ # minimal theme
  scale_x_discrete(position = "top") +
  # labs(x=expression(italic("S. thermophilus")),y=expression(italic("L. delbrueckii")))+
 theme(axis.text.y = element_text(size=7),
      axis.text.x = element_text(angle = 45,vjust=1,
    size = 12, hjust = 0),axis.title=element_blank(), 
    legend.title = element_blank())
 # coord_fixed()
# Print the heatmap
print(ggheatmap)

##----------------------------------
#count mechanisms

unique(pathway_identification_species$mechanism_final)

sum(pathway_identification_species$Present)

StrainCasType_generel <- StrainCasType %>% filter(mechanism_final=="CRISPR") 
table(StrainCasType_generel$species)

StrainCasType_nongenerel <- StrainCasType %>% filter(mechanism_final!="CRISPR") 
tmp <- StrainCasType_nongenerel %>% select(-mechanism_final) %>% group_by(species,Strain) %>% summarise(countsss=n())
table(tmp$countsss)

(121+49+1)/831

tmp2 <- tmp %>% filter(countsss>1)
table(tmp2$species)/(121+49+1) %>% sort()

##======================================================
###--------------------------make discrete
##======================================================

ggplot(Strain_all_Type_final,aes(x=percent_species))+theme_classic()+geom_histogram()+geom_vline(xintercept = c(10,90))

Strain_all_Type_final$category <- ifelse(Strain_all_Type_final$percent_species==0,"NA",ifelse(Strain_all_Type_final$percent_species<10,"low",ifelse(Strain_all_Type_final$percent_species<90,"intermediate","high")))

Strain_all_Type_final$category = factor(Strain_all_Type_final$category, levels=c("NA","low","intermediate","high"))

colorsss <- rev(c("#E72405","#FB9B19","#F7F808","white"))

ggheatmap_02 <- ggplot(Strain_all_Type_final, aes(y=species, x=mechanism_final, fill = category))+
 geom_tile(color = "white",size=1.1)+
 scale_fill_manual(values = colorsss)+
  theme_classic()+
  scale_x_discrete(position = "top") +
 theme(axis.text.y = element_text(size=7),
      axis.text.x = element_text(angle = 45,vjust=1,
    size = 12, hjust = 0),axis.title=element_blank(), 
    legend.title = element_blank())

# Print the heatmap
print(ggheatmap_02)

Strain_all_Type_final_sub <- Strain_all_Type_final %>% filter(category %in% c("low","intermediate","high")) 

colorsssss <- rev(c("#E72405","#FB9B19","#F7F808"))

ggplot(Strain_all_Type_final_sub,aes(x=species,fill=category))+theme_classic()+geom_bar()+ scale_fill_manual(values = colorsssss)+coord_flip()

```

## core, accessory counts 

Here, I visulise the absolut and relative abundance of core, accesory and cloud defense mechanisms. 

```{r }
####--------------------------------------------
#add number of core,accessory and cloud immune systems
####--------------------------------------------

Strain_all_Type_final_sub_01 <- read_csv("data_submit/Strain_all_Type_final_sub_data.csv")

colorsssss <- rev(c("#E72405","#FB9B19","#F7F808"))

ggplot(Strain_all_Type_final_sub_01,aes(x=species,fill=category))+theme_classic()+geom_bar()+ scale_fill_manual(values = colorsssss)

Braod_species_tree_raw$tip.label
unique(Strain_all_Type_final_sub_01$species)

Strain_all_Type_final_sub_01_prep <- Strain_all_Type_final_sub_01 %>% group_by(species,category) %>% summarise(value=n())

Strain_all_Type_final_sub_01_prep$category = factor(Strain_all_Type_final_sub_01_prep$category, levels=c("low","intermediate","high"))

p12 <- gtree_01 + geom_facet(panel = 'bar', data = Strain_all_Type_final_sub_01_prep, geom = geom_bar, 
                 mapping = aes(x = value, fill = as.factor(category)), 
                 orientation = 'y', width = 0.8, stat='identity') +scale_fill_manual(values =colorsssss )

facet_widths(p12, widths = c(1, 2))

# pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/pan_immunity_count.pdf",width=10,height=6)

# facet_widths(p12, widths = c(1, 2))

# dev.off()

Strain_all_Type_final_sub_01_prep_02 <- Strain_all_Type_final_sub_01 %>% group_by(species,category) %>% summarise(ind=n()) %>% group_by(species) %>% summarise(total=sum(ind)) %>% left_join(.,Strain_all_Type_final_sub_01_prep) %>% mutate(percent=100*(value/total))

Strain_all_Type_final_sub_01_prep$category = factor(Strain_all_Type_final_sub_01_prep$category, levels=c("low","intermediate","high"))

p13 <- gtree_01 + geom_facet(panel = 'bar', data = Strain_all_Type_final_sub_01_prep_02, geom = geom_bar, 
                 mapping = aes(x = percent, fill = as.factor(category)), 
                 orientation = 'y', width = 0.8, stat='identity') +scale_fill_manual(values =colorsssss )

facet_widths(p13, widths = c(1, 2))


   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/pan_immunity_frequency.pdf",width=10,height=6)

p13
# dev.off()
```


## look at non-CRISPR species

```{r}
c("Brevibacterium_aurantiacum", "Brevibacterium_linens", "Companilactobacillus_versmoldensis", "Lactococcus_lactis","Leuconostoc_mesenteroides")
dim(pathway_identification_species)
sum(pathway_identification_species$Present)
pathway_identification_species$grouping <- ifelse(pathway_identification_species$species %in% c("Brevibacterium_aurantiacum", "Brevibacterium_linens", "Companilactobacillus_versmoldensis", "Lactococcus_lactis","Leuconostoc_mesenteroides"),"No CRISPR","CRISPR")

crisprss <- pathway_identification_species %>% filter(percent_species>0) %>% ungroup()%>% group_by(species,grouping) %>% dplyr::summarise(systems=n())

crisprss

res <- wilcox.test(systems ~ grouping, data = crisprss)
res

women_weight <- crisprss %>% filter(grouping=="CRISPR") %>% ungroup() %>% select(systems) %>% unlist() %>% as.vector()
men_weight <- crisprss %>% filter(grouping!="CRISPR") %>% ungroup() %>% select(systems) %>% unlist()%>% as.vector()
res <- wilcox.test(women_weight, men_weight)
res$p.value

t.test(women_weight,men_weight)

phage_defense_plot <- ggplot(crisprss,aes(x=grouping,y=systems))+geom_boxplot()+theme_classic()+labs(x="",y="Phage defense systems",title = "species-level")#+geom_jitter()
phage_defense_plot


phage_defense_plot <- ggplot(crisprss,aes(x=grouping,y=systems,fill=grouping,color=grouping))+geom_violin(alpha=0.6,adjust=0.8)+theme_classic()+labs(x="",y="Phage defense systems")+ggforce::geom_sina(alpha=0.8)+labs(x="",y="Phage defense systems",title = "species-level")+ stat_compare_means()

phage_defense_plot

  # png("~/Desktop/Projects/2020_CRISPRscope/Plots/phage_defense_system_count.png", width = 1200, height = 1500,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)
phage_defense_plot
 # dev.off()
 
 ###----------------------------------------
 #per strain
  ###----------------------------------------
 
 
```

do this per strain

```{r}

pathway_identification
table(pathway_identification$mechanism_final)

StrainCasType_tmp <- StrainCasType %>% filter(mechanism_final=="CRISPR")

StrainCasType_tmp_02 <- StrainCasType %>% mutate(genome=paste0(species,"_",Strain,".faa")) %>% ungroup() %>% select(genome,mechanism_final) %>% filter(mechanism_final=="CRISPR")
pathway_identification_tmp_02 <- pathway_identification %>% filter(Completness=="Yes")%>% ungroup() %>% select(genome,mechanism_final)

# merge(pathway_identification,StrainCasType_tmp,by=c(""))

all_defense <- rbind(pathway_identification_tmp_02,StrainCasType_tmp_02)  %>% add_column(all=1)%>% spread(., mechanism_final, all) %>%replace(is.na(.), all_defense0) 

colnames(all_defense)
ncol(all_defense)


all_defense$numDefense <- rowSums(all_defense[,c(2:4,6:24)])

all_defense$CRISPR_presence <- ifelse(all_defense$CRISPR=="0","No CRISPR-containing\nstrains","CRISPR-containing\nstrains")

all_defense_crisprsss <- all_defense %>% filter(!species %in% c("Brevibacterium_aurantiacum", "Brevibacterium_linens", "Companilactobacillus_versmoldensis", "Lactococcus_lactis","Leuconostoc_mesenteroides"))

# phage_defense_plot_strain_02 <- ggplot(all_defense,aes(x=CRISPR_presence,y=numDefense))+geom_boxplot()+theme_classic()+labs(x="",y="Phage defense systems",title = "strain-level")+ stat_compare_means()

phage_defense_plot_strain_02 <- ggplot(all_defense,aes(x=CRISPR_presence,y=numDefense,fill=CRISPR_presence,color=CRISPR_presence))+geom_violin(alpha=0.6,adjust=1.7)+theme_classic()+labs(x="",y="Phage defense systems")+ggforce::geom_sina(alpha=0.2)+labs(x="",y="Phage defense systems",title = "strain-level")+ stat_compare_means()+geom_boxplot(color="black",alpha=0,width=.25)

phage_defense_plot_strain_02

  # png("~/Desktop/Projects/2020_CRISPRscope/Plots/phage_defense_system_count_species_strains.png", width = 2400, height = 1500,res=300)
     # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//phage_defense_system_count_species_strains.pdf",width=,height=4)

   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)
(phage_defense_plot+theme(legend.position = "none"))+(phage_defense_plot_strain_02+theme(legend.position = "none"))
 # dev.off()
 
  ###----------------------------------------
 #per species 
  ###----------------------------------------
 
 #add species info
 
 all_defense$Genus <- str_split_fixed(all_defense$genome, fixed("_"), 3)[,1]
all_defense$spec <- str_split_fixed(all_defense$genome, fixed("_"), 3)[,2]
all_defense$species <- paste(all_defense$Genus,all_defense$spec,sep="_")
 
table(all_defense$species)


 all_defense$grouping <- ifelse(all_defense$species %in% c("Brevibacterium_aurantiacum", "Brevibacterium_linens", "Companilactobacillus_versmoldensis", "Lactococcus_lactis","Leuconostoc_mesenteroides"),"No CRISPR-containing\nspecies","CRISPR-containing\nspecies")


 phage_defense_plot_species <- ggplot(all_defense,aes(x=grouping,y=numDefense))+geom_boxplot()+theme_classic()+labs(x="",y="Phage defense systems",title = "species-level")+ stat_compare_means()#+geom_jitter(alpha=0.1)

phage_defense_plot_species


phage_defense_plot_species <- ggplot(all_defense,aes(x=grouping,y=numDefense,fill=grouping,color=grouping))+geom_violin(alpha=0.6,adjust=1.7)+theme_classic()+labs(x="",y="Phage defense systems")+ggforce::geom_sina(alpha=0.2)+labs(x="",y="Phage defense systems",title = "species-level")+ stat_compare_means()+geom_boxplot(color="black",alpha=0,width=.25)

phage_defense_plot_species

phage_defense_plot_species+phage_defense_plot_strain_02
 (phage_defense_plot_species+theme(legend.position = "none"))+(phage_defense_plot_strain_02+theme(legend.position = "none"))

  # png("~/Desktop/Projects/2020_CRISPRscope/Plots/phage_defense_system_count_species_strains.png", width = 2400, height = 1500,res=300)
   pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//phage_defense_system_count_species_strains.pdf",width=,height=3.7)
    (phage_defense_plot_species+theme(legend.position = "none"))+(phage_defense_plot_strain_02+theme(legend.position = "none"))

# phage_defense_plot+phage_defense_plot_strain_02
 dev.off()
  
```

## PCA defense mechanisms

Here, I try to cluster the the strains according to the defenese mechanisms presence

```{r}
 defense_systems <-  read_csv("data_submit/pathway_strain.txt") %>% filter(Completness=="Yes")%>% select(species,genome,mechanism_final)


StrainCasType_subtypes <- read_csv("data_submit/StrainCasType_withSubtype.csv") %>% rename(., "species"="Organism")%>% mutate(mechanism_final=paste0("CRISPR type ",Type,"_",Subtype))%>% filter(mechanism_final!="CRISPR type III_B")  %>% mutate(genome=paste0(species,"_",Strain,".faa"))  %>% select(species,genome,mechanism_final)


defense_systems_complete <- rbind(defense_systems,StrainCasType_subtypes) %>% mutate(presence=1)

defense_systems_complete <- rbind(defense_systems,StrainCasType_subtypes) %>% mutate(presence=1) %>% filter(!species %in% c("Companilactobacillus_versmoldensis","Pediococcus_parvulus","Corynebacterium_casei","Brevibacterium_linens","Loigolactobacillus_coryniformis"))
#,"Mammaliicoccus_sciuri","Staphylococcus equorum"

  defense_systems_complete_wide <- spread(defense_systems_complete, mechanism_final, presence)%>%replace(is.na(.), 0) 
table(defense_systems_complete_wide$species)
  defense_systems_species <- defense_systems_complete_wide %>% select(genome,species)

  
defense_systems_complete_wide_preped <- defense_systems_complete_wide %>% select(-species) %>% remove_rownames()%>% column_to_rownames(., var = "genome")



df_pca <- prcomp(defense_systems_complete_wide_preped)

plot(df_pca$x[,1], df_pca$x[,2])

metabolo.sum <- summary(df_pca)
metabolo.sum
metabolo.sum <- metabolo.sum$importance %>% as.data.frame()

ploting_data <- df_pca$x %>% as.data.frame()
ploting_data$genome <- rownames(ploting_data)
ploting_data_final <-  merge(ploting_data,defense_systems_species,by="genome")

table(ploting_data_final$species)

pca0 <- ggbiplot(df_pca,var.axes=FALSE,obs.scale = 1)+theme_classic()+labs(title="grouping by biological replicates")+theme(legend.position = "none")
pca0

pca1 <- ggbiplot(df_pca,var.axes=FALSE,groups=ploting_data_final$species,obs.scale = 1,alpha=0.3,ellipse=TRUE)+theme_classic()+theme(legend.position = "right")
pca1

species_colors <- read_delim("data_submit/species_colors.txt", "\t", escape_double = FALSE, trim_ws = TRUE)

###---------------following names are used in IMG
unique(species_colors$species)
species_colors$species <- plyr::revalue(species_colors$species,c(
 "Propionibacterium acnes"="Cutibacterium acnes"  , 
"Lactobacillus plantarum" ="Lactiplantibacillus plantarum",
# "Lactobacillus_versmoldensis"     = "Companilactobacillus_versmoldensis",
"Lactobacillus curvatus"          = "Latilactobacillus curvatus",
 "Lactobacillus rhamnosus"         = "Lacticaseibacillus rhamnosus",
# "Lactobacillus_coryniformis"      = "Loigolactobacillus_coryniformis",
 # "Lactobacillus_rhamnosus"         = "Mammaliicoccus sciuri",
 "Streptococcus luteciae"         = "Streptococcus lutetiensis",
"Lactobacillus fermentum"         = "Limosilactobacillus fermentum",
 "Lactobacillus zeae"              = "Lacticaseibacillus casei"))
unique(species_colors$species)

table(species_colors$species)

ploting_data_final$species_new <- gsub("_"," ",ploting_data_final$species)


unique(species_colors$species)
unique(ploting_data_final$species)
unique(ploting_data_final$species_new)

nrow(species_colors)
unique(ploting_data_final$species_new) %>% length()

species_colors_02 <- species_colors %>% filter(species %in% unique(ploting_data_final$species_new) )
nrow(species_colors_02)
unique(ploting_data_final$species_new) %>% length()

unique(ploting_data_final$species_new)[which(!unique(ploting_data_final$species_new) %in% species_colors_02$species)]

species_colors_03 <- rbind(species_colors_02,c("Mammaliicoccus sciuri","grey"))



target <- unique(ploting_data_final$species_new) %>% as.vector

table(ploting_data_final$species_new)

species_colors_03 <- species_colors_03[match(target, species_colors_03$species),]
ploting_data_final$species_new = factor(ploting_data_final$species_new, levels=species_colors_03$species)

pca2 <- ggbiplot(df_pca,var.axes=FALSE,groups=ploting_data_final$species_new,obs.scale = 1,alpha=0.3,size=1.2,ellipse=TRUE)+theme_classic()+theme(legend.position = "right")+scale_fill_manual(values =species_colors_03$color )+scale_color_manual(values =species_colors_03$color )
pca2

pca3 <- ggbiplot(df_pca,var.axes=FALSE,groups=ploting_data_final$species_new,obs.scale = 1,alpha=0.3,size=1.2,ellipse=TRUE)+theme_classic()+theme(legend.position = "none")+scale_fill_manual(values =species_colors_03$color )+scale_color_manual(values =species_colors_03$color )
pca3

  # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//pca_defense_systems.pdf",width=6,height=3)
# pca3
 # dev.off()


# ##-----------tsne
# library(Rtsne)
# tsne.norm = Rtsne(pca.norm$x, pca = FALSE)
# tsne_plot <- as.data.frame(tsne.norm$Y)
# tsne_plot$site <- sample_relative_safe_final_tsne_safe_sterm_woDuplicates[,1]
# tsne_plot$contig <- paste(str_split_fixed(tsne_plot$site, "_", 5)[,1],str_split_fixed(tsne_plot$site, "_", 5)[,2],sep = "_")
# 
# 
# tsne_plot$species <- tsne_plot$contig
# 
# ##plot blank
# ggplot(tsne_plot,aes(x=V1,y=V2))+geom_point()

```

## within species defense decrease

run previous chunk before
```{r}
defense_systems_complete_wide


 ANI_CUSTOM_OUTPUT_21072021 <- read_csv("data_submit/ANI_perArray_sharedSpacers_newer.csv") 

table(ANI_CUSTOM_OUTPUT_21072021$Organism1)

ANI_CUSTOM_OUTPUT_21072021 <- read_csv("data_submit/ANI_CUSTOM_OUTPUT_21072021.csv") %>% mutate(genome1=strain_query,genome2=strain_ref) %>% select(genome1,genome2,ANI)

fastani_thibualt <- read_delim("data_submit/20201219_ANI_output_all.txt", "\t", escape_double = FALSE, col_names = FALSE,  trim_ws = TRUE)

fastani_thibualt$genome1 <- str_split_fixed(fastani_thibualt$X1, fixed("/"), 9)[,8] %>%  gsub(".fna","",.)
fastani_thibualt$genome2 <- str_split_fixed(fastani_thibualt$X2, fixed("/"), 9)[,8] %>%  gsub(".fna","",.)

fastani_thibualt_cleaned <- fastani_thibualt %>% mutate(ANI=X3) %>% select(c(genome1,genome2,ANI="X3"))

defense_systems_complete_wide <- defense_systems_complete_wide %>% mutate(strain=gsub(".*_","",genome))  %>% mutate(strain=gsub(".faa$","",strain))

table(defense_systems_complete_wide$species)


fastani_thibualt_cleaned_total <- rbind(ANI_CUSTOM_OUTPUT_21072021,fastani_thibualt_cleaned) %>% unique()


speciesss="Bifidobacterium_animalis"
strainName_1="GCF.000025245.2"
strainName_2="GCF.000025245.2"



defense_systems_complete_wide$strain %in% fastani_thibualt_cleaned$genome1 %>% length()
defense_systems_complete_wide$strain %in% fastani_thibualt_cleaned$genome1 %>% sum()
defense_systems_complete_wide_avail <- defense_systems_complete_wide[(defense_systems_complete_wide$strain %in% fastani_thibualt_cleaned_total$genome2),]

defense_systems_complete_wide_avail$strain %in% fastani_thibualt_cleaned_total$genome1%>% length()

# table(defense_systems_complete_wide_avail$species)^2
table(defense_systems_complete_wide_avail$species)

raw_Output <- data.frame()
final_Output <- data.frame()

for (speciesss in unique(defense_systems_complete_wide_avail$species)) {
  print(speciesss)
  
  tmp1 <- defense_systems_complete_wide_avail %>% filter(species==speciesss)
  
  numGen=unique(tmp1$strain) %>% length()
  # counter_1=1
  counter_2=unique(tmp1$strain) %>% length()
  
  # ifelse(numGen<80,vector=1:numGen, sample(1:numGen, 80))
  if (numGen<30) {
    vector <- 1:numGen
  }   else {
vector <- sample(1:numGen, 30)}
 
  
  for (strains1 in vector) {
    
    
    strainName_1=unique(tmp1$strain)[strains1]
    
    for (strains2 in 1:counter_2) {
      
          strainName_2=unique(tmp1$strain)[strains2]
      # counter_2=counter_2-1

     
    
      
      ##----analysis
      
     tmp2 <- tmp1 %>% filter(strain %in% c(strainName_1,strainName_2)) %>% select(-species,-genome,-strain) %>% colSums() %>% table() %>% data.frame() %>% rename(c("matching"=".","abundance"="Freq")) %>% mutate(species=speciesss,strain1=strainName_1,strain2=strainName_2)
     raw_Output <- rbind(raw_Output,tmp2)
     
     tmp3 <- tmp2 %>% filter(matching %in% c(0,2)) %>% pull(abundance) %>% sum()
      
# tmp5 <-  ANI_CUSTOM_OUTPUT_21072021 %>% filter((Strain1==strainName_1 & Strain2==strainName_2) | (Strain1==strainName_2 & Strain2==strainName_1) )   %>% pull(ANI) %>% median()

 # fastani_thibualt_cleaned_total %>% filter((genome1==strainName_1))
     tmp5 <- fastani_thibualt_cleaned_total %>% filter((genome1==strainName_1 & genome2==strainName_2) | (genome1==strainName_2 & genome2==strainName_1)) %>% pull(ANI) %>% median()

    # tmp5 <- ANI_CUSTOM_OUTPUT_21072021 %>% filter((strain_query==strainName_1 & strain_ref==strainName_2) | (strain_query==strainName_2 & strain_ref==strainName_1)) %>% pull(ANI) %>% median()

               tmp4 <- data.frame(species=speciesss,strain1=strainName_1,strain2=strainName_2,numberMatching=tmp3,percentMatch=100*(tmp3/29),ANI=tmp5)
     final_Output  <- rbind(final_Output,tmp4)
     print(tmp4)
  }
    
     counter_2=(counter_2-1)
  }
  
  
}


  # write.csv(final_Output,file="~/Desktop/Projects/2020_CRISPRscope/data/fastani_immunesystem_correlation_01.csv")

  

final_Output %>% nrow()
final_Output %>% filter(!is.na(ANI))%>% nrow()

ANI_CUSTOM_OUTPUT_21072021 %>% filter(species_query==speciesss) %>% pull(strain_query) %>% table() %>% length()
```


```{r}
###--------------------------------------------------plot
final_Output_cleaned <- read_csv("data_submit/fastani_immunesystem_correlation_01.csv")


final_Output_cleaned_sub <- final_Output_cleaned %>% filter(ANI>97.5)

ggplot(final_Output_cleaned_sub,aes(x=ANI,y=percentMatch))+geom_point(alpha=0.1)+facet_wrap(~species,scales = "free")+theme_classic()+geom_smooth(method='lm')


ggplot(final_Output_cleaned_sub,aes(x=ANI,y=percentMatch))+geom_point(alpha=0.1)+theme_classic()+geom_smooth(method='lm')


# ggplot(final_Output_cleaned_sub,aes(y=ANI,x=as.character(percentMatch)))+geom_boxplot()+facet_wrap(~species)+theme_classic() + theme(axis.text.x = element_text(angle = 45,vjust=1))
    


###--------------------------------------------------violin

final_Output_cleaned_sub_above <- final_Output_cleaned_sub %>% filter(percentMatch>75) %>% mutate(numberDismatching=29-numberMatching)%>% filter(numberDismatching<6)

ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(percentMatch)))+geom_point(alpha=0.1)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))

ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(percentMatch)))+geom_violin()+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))
ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(percentMatch)))+geom_boxplot()+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))

ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(percentMatch)))+geom_jitter(alpha=0.1)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')


ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(numberDismatching)))+geom_jitter(alpha=0.1)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')

ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(numberDismatching)))+geom_jitter(alpha=0.1)+geom_violin()+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')


ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(numberDismatching)))+geom_jitter(alpha=0.1)+geom_violin()+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')+facet_wrap(~species)
table(final_Output_cleaned_sub_above$numberDismatching)

ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=as.factor(numberDismatching)))+geom_jitter(alpha=0.1)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')+facet_wrap(~species)

ggplot(final_Output_cleaned_sub_above,aes(y=ANI,x=(numberDismatching)))+geom_jitter(alpha=0.1)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')+facet_wrap(~species)


species_of_interest <- table(final_Output_cleaned_sub_above$species) %>% data.frame() %>% filter(Freq>500) %>% filter(!Var1 %in% c("Bifidobacterium_animalis", "Lactobacillus_acidophilus"))%>% pull(Var1) %>% as.character() 

final_Output_cleaned_sub_above_dom <- final_Output_cleaned_sub_above %>% filter(species %in% species_of_interest)

ggplot(final_Output_cleaned_sub_above_dom,aes(y=ANI,x=(numberDismatching)))+geom_jitter(alpha=0.1)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')+facet_wrap(~species)
ggplot(final_Output_cleaned_sub_above_dom,aes(y=ANI,x=(numberDismatching)))+geom_jitter(alpha=0.05)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+geom_smooth(method='lm')


##---------------------------------------------------
#split
##---------------------------------------------------
my.formula <- y ~ x
library(ggpmisc)
plot <- ggplot(final_Output_cleaned_sub_above_dom,aes(x=numberDismatching,y=ANI))+geom_jitter(alpha=0.05)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+
    scale_y_continuous(limit=c(97.5,100))+
geom_smooth(method="lm",se = FALSE,color="blue",linetype = "dashed")+
  stat_poly_eq(formula = my.formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               parse = TRUE,)


plot

model <- lm(ANI ~ numberDismatching, data = final_Output_cleaned_sub_above_dom)
model
summary(model)
#-------------------important subtypes
###suptypes 

CRISPR_spacer_coverage_filtered_subs <- CRISPR_spacer_coverage_filtered %>% filter(Subtype %in% c("I-E","I-C","II-C","I-G","II-A","III-A"))

  
plot2 <- ggplot(final_Output_cleaned_sub_above_dom,aes(x=numberDismatching,y=ANI,color=species))+geom_jitter(alpha=0.05)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+
# coord_trans(y="log2",x="log2")+
    # scale_x_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  scale_y_continuous(limit=c(97.5,100))+
  # lims(x=c(0,100000),y=c(0,100000))+
  # labs(x="Spacer (cpm)",y="Protospacers  (cpm)")+
geom_smooth(method="lm",se = FALSE)

plot2

plot+plot2

colorDataFrame <- data.frame(species= c( "Lactobacillus acidophilus",
"Lactobacillus fermentum", "Pediococcus parvulus",
"Bifidobacterium animalis", "Brevibacterium aurantiacum", "Lactobacillus versmoldensis",
"Leuconostoc lactis", "Corynebacterium casei", "Pediococcus pentosaceus",
"Propionibacterium freudenreichii", "Streptococcus luteciae",
"Lactobacillus rhamnosus", "Lactobacillus coryniformis", "Lactobacillus curvatus",
"Brevibacterium linens", "Leuconostoc pseudomesenteroides", "Lactobacillus plantarum",
"Staphylococcus sciuri", "Leuconostoc mesenteroides", "Propionibacterium acnes",
"Staphylococcus equorum", "Lactobacillus casei", "Lactococcus lactis",
"Lactobacillus helveticus", "Lactobacillus delbrueckii", "Streptococcus thermophilus"
),
color= c("#fdc2b496","#a4778bff","#a3758a64","#17c3b2ff","#16c3b396","#a7c5da4b","#ff6d584b","grey","#a3768aaf","#61988eff","#007ea796","#a7c5da96","#eeabc4ff","#eeaac396","#14c2b14b","#ff6f59ff","#a3758a32","#c8beb7ff","#ff6e5896","grey","grey","#a7bed3ff","#eeba0bff","#fcc2b44b","#fdc2b5ff","#007fa6ff"))
# # 17c3b2ff

colorDataFrame$species <- plyr::revalue(colorDataFrame$species,c(
 "Propionibacterium acnes"="Cutibacterium acnes"  , 
"Lactobacillus plantarum" ="Lactiplantibacillus plantarum",
# "Lactobacillus_versmoldensis"     = "Companilactobacillus_versmoldensis",
"Lactobacillus curvatus"          = "Latilactobacillus curvatus",
 "Lactobacillus rhamnosus"         = "Lacticaseibacillus rhamnosus",
# "Lactobacillus_coryniformis"      = "Loigolactobacillus_coryniformis",
 # "Lactobacillus_rhamnosus"         = "Mammaliicoccus sciuri",
 "Streptococcus luteciae"         = "Streptococcus lutetiensis",
"Lactobacillus fermentum"         = "Limosilactobacillus fermentum",
 "Lactobacillus zeae"              = "Lacticaseibacillus casei"))
# unique(species_colors$species)

colorDataFrame$species <- gsub(" ","_",colorDataFrame$species)
colorDataFrame_interest <-colorDataFrame %>% filter(species %in% species_of_interest)

colorDataFrame_interest$species = factor(colorDataFrame_interest$species, levels=colorDataFrame_interest$species)


plot2 <- ggplot(final_Output_cleaned_sub_above_dom,aes(x=numberDismatching,y=ANI,color=species))+geom_jitter(alpha=0.05)+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+
# coord_trans(y="log2",x="log2")+
    scale_x_continuous(breaks=c(0,1,2,3,4,5))+
  scale_color_manual(values = colorDataFrame_interest$color)+
  scale_y_continuous(limit=c(97.5,100))+
  # lims(x=c(0,100000),y=c(0,100000))+
  labs(y="ANI (%)",x="Number of different\naccessory immunity systems")+
geom_smooth(method="lm",se = FALSE)

plot2


##---------------------------------------------------
#split
##---------------------------------------------------


plot


   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//decreasing_ANI_accessory.pdf",width=6,height=4)
# plot2+theme(legend.position = "none")
 # dev.off()

 
 plot3 <- ggplot(final_Output_cleaned_sub_above_dom,aes(x=numberDismatching,y=ANI,color=species))+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+
# coord_trans(y="log2",x="log2")+
    scale_x_continuous(breaks=c(0,1,2,3,4,5))+
  scale_color_manual(values = colorDataFrame_interest$color)+
  scale_y_continuous(limit=c(97.5,100))+
  # lims(x=c(0,100000),y=c(0,100000))+
  labs(y="ANI (%)",x="Number of different\naccessory immunity systems")+
geom_smooth(method="lm",se = FALSE)+
  stat_poly_eq(formula = my.formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),label.x=9,
               parse = TRUE,)

plot3
 # label.x=7,
# +geom_jitter(alpha=0.05)
 
     # png("~/Desktop/Projects/2020_CRISPRscope/Plots//decreasing_ANI_accessory_supp.png", width = 2100, height = 1500,res=300)

   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//decreasing_ANI_accessory_supp.pdf",width=6,height=4)
# plot3

# dev.off()
```

# Figure 2



## ANI vs. CRISPR

```{r}
library(tidyverse)
ANI_CUSTOM_OUTPUT_21072021 <- read_csv("data_submit/ANI_CUSTOM_OUTPUT_21072021.csv") %>% filter(turnover_rate<Inf) %>% filter(perc_shared!=0)

table(ANI_CUSTOM_OUTPUT_21072021$species_query)

ANI_CUSTOM_OUTPUT_21072021 %>% filter(perc_shared>0.5) %>% summarise(median=median(ANI),means=mean(ANI),sds=sd(ANI))

ggplot(ANI_CUSTOM_OUTPUT_21072021,aes(x=ANI,perc_shared))+geom_point(alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~species_ref)+coord_trans(y="log10")


ggplot(ANI_CUSTOM_OUTPUT_21072021,aes(x=ANI,perc_shared))+geom_point(alpha=0.1,size=0.2)+theme_classic()+coord_trans(y="log10")

ANI_CUSTOM_OUTPUT_21072021 %>% filter(ANI>99.5) %>% summarise(median=100*median(perc_shared),means=100*mean(perc_shared),sds=100*sd(perc_shared))

ANI_CUSTOM_OUTPUT_21072021 %>% filter(ANI>99.9) %>% summarise(median=100*median(perc_shared),means=100*mean(perc_shared),sds=100*sd(perc_shared))


ANI_CUSTOM_OUTPUT_21072021 %>% filter(ANI<99) %>% summarise(median=100*median(perc_shared),means=100*mean(perc_shared),sds=100*sd(perc_shared))


# %>% filter(species_ref=="Lacticaseibacillus_casei") 
##-----------------
ANI_CUSTOM_OUTPUT_21072021$ANI_rounded <- round(ANI_CUSTOM_OUTPUT_21072021$ANI,digits = 1)


ggplot(ANI_CUSTOM_OUTPUT_21072021,aes(group=ANI_rounded,x=ANI_rounded,perc_shared))+geom_boxplot(alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~species_ref)+coord_trans(y="log10")+geom_point(alpha=0.1,size=0.2)


numbercomparison_df <- ANI_CUSTOM_OUTPUT_21072021 %>% group_by(species_ref,ANI_rounded) %>% summarise(numbercomparisons_tmp=n())

ggplot(numbercomparison_df,aes(numbercomparisons_tmp))+geom_bar()+theme_classic()+coord_trans(x="log10")


cutoff_n_comparison=40
# numbercomparison_df$numbercomparisons_tmp>
keeps <- numbercomparison_df %>% filter(numbercomparisons_tmp>cutoff_n_comparison)

ANI_CUSTOM_OUTPUT_21072021_keeps <- data.frame()

for (numRows in 1:nrow(keeps)) {
  speciessss <- keeps[numRows,"species_ref"] %>% unlist() %>% as.character()
  ANIsss <- keeps[numRows,"ANI_rounded"]%>% unlist() %>% as.character()
# ANI_CUSTOM_OUTPUT_21072021$species_ref==speciessss
  # tmpss <- ANI_CUSTOM_OUTPUT_21072021 %>% filter(species_ref==speciessss)
  tmpss <- ANI_CUSTOM_OUTPUT_21072021 %>% filter(species_ref==speciessss&ANI_rounded==ANIsss)

  ANI_CUSTOM_OUTPUT_21072021_keeps <- rbind(ANI_CUSTOM_OUTPUT_21072021_keeps,tmpss)
}



plot_ANI_Percent_shared_species <- ggplot(ANI_CUSTOM_OUTPUT_21072021_keeps,aes(group=ANI_rounded,x=ANI_rounded,perc_shared))+geom_boxplot(outlier.shape = NA,alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~species_ref)+coord_trans(y="log10")#+geom_point(alpha=0.1,size=0.2)

unique(ANI_CUSTOM_OUTPUT_21072021_keeps$species_ref)
ANI_CUSTOM_OUTPUT_21072021_keeps_cleaned <- ANI_CUSTOM_OUTPUT_21072021_keeps %>% filter(species_ref %in% c("Lactiplantibacillus_plantarum","Lactobacillus_delbrueckii", "Lactobacillus_helveticus","Propionibacterium_freudenreichii","Streptococcus_thermophilus"))

plot_ANI_Percent_shared_species <- ggplot(ANI_CUSTOM_OUTPUT_21072021_keeps_cleaned,aes(group=ANI_rounded,x=ANI_rounded,y=100*(perc_shared)))+geom_boxplot(outlier.shape = NA,alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~species_ref)+coord_trans(y="log10")+labs(x="ANI (%)",y="shared CRISPR (%)")
plot_ANI_Percent_shared_species

  # png("~/Desktop/Projects/2020_CRISPRscope/Plots//CRISPRshared_ANI_perSpecies.png", width = 2100, height = 1500,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)

# plot_ANI_Percent_shared_species
 # dev.off()

##-----------------only above 99.5

ANI_CUSTOM_OUTPUT_21072021_keeps_highs <- ANI_CUSTOM_OUTPUT_21072021_keeps %>% filter(ANI_rounded>99.4)

ggplot(ANI_CUSTOM_OUTPUT_21072021_keeps_highs,aes(group=ANI_rounded,x=ANI_rounded,perc_shared))+geom_boxplot(outlier.shape = NA,alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~species_ref)+coord_trans(y="log10")#+geom_point(alpha=0.1,size=0.2)


```

new ANI vs. shared CRISPR for arrays

```{r}
library(tidyverse)

 ANI_perArray_sharedSpacers_new <- read_csv("data_submit/ANI_perArray_sharedSpacers_newer.csv") %>% mutate(shared_spacers=100*perc_shared) %>% filter(shared_spacers!=0) %>% filter(Cas_subtype1==Cas_subtype2)
table(ANI_perArray_sharedSpacers_new$Organism1)

ggplot(ANI_perArray_sharedSpacers_new,aes(x=ANI,y=shared_spacers))+geom_point(alpha=0.2,size=0.4)+theme_classic()+coord_trans(y="log2")
ggplot(ANI_perArray_sharedSpacers_new,aes(x=ANI,y=shared_spacers))+geom_point(alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Organism1)+coord_trans(y="log10")

ggplot(ANI_perArray_sharedSpacers_new,aes(x=ANI,y=shared_spacers))+geom_point(alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Cas_subtype1)+coord_trans(y="log10")


ggplot(ANI_perArray_sharedSpacers_new,aes(x=ANI))+geom_density()+theme_classic()
ggplot(ANI_perArray_sharedSpacers_new,aes(x=shared_spacers))+geom_density()+theme_classic()+coord_trans(x="log10")


ANI_perArray_sharedSpacers_new$shared_spacers
ANI_perArray_sharedSpacers_new %>% filter(ANI>99.5) %>% summarise(median=median(shared_spacers),means=mean(shared_spacers),sds=sd(shared_spacers))

##-----------------
ANI_perArray_sharedSpacers_new$ANI_rounded <- round(ANI_perArray_sharedSpacers_new$ANI,digits = 1)


ggplot(ANI_perArray_sharedSpacers_new,aes(group=ANI_rounded,x=ANI_rounded,shared_spacers))+geom_boxplot(alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Organism1)+coord_trans(y="log10")+geom_point(alpha=0.1,size=0.2)
ggplot(ANI_perArray_sharedSpacers_new,aes(group=ANI_rounded,x=ANI_rounded,shared_spacers))+geom_boxplot(alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Cas_subtype1)+coord_trans(y="log10")+geom_point(alpha=0.1,size=0.2)

numbercomparison_df <- ANI_perArray_sharedSpacers_new %>% group_by(Cas_subtype1,ANI_rounded) %>% summarise(numbercomparisons_tmp=n())

ggplot(numbercomparison_df,aes(numbercomparisons_tmp))+geom_bar()+theme_classic()+coord_trans(x="log10")


cutoff_n_comparison=40
# numbercomparison_df$numbercomparisons_tmp>
keeps <- numbercomparison_df %>% filter(numbercomparisons_tmp>cutoff_n_comparison)

ANI_CUSTOM_OUTPUT_21072021_keeps <- data.frame()

for (numRows in 1:nrow(keeps)) {
  speciessss <- keeps[numRows,"Cas_subtype1"] %>% unlist() %>% as.character()
  ANIsss <- keeps[numRows,"ANI_rounded"]%>% unlist() %>% as.character()
# ANI_CUSTOM_OUTPUT_21072021$species_ref==speciessss
  # tmpss <- ANI_CUSTOM_OUTPUT_21072021 %>% filter(species_ref==speciessss)
  tmpss <- ANI_perArray_sharedSpacers_new %>% filter(Cas_subtype1==speciessss&ANI_rounded==ANIsss)

  ANI_CUSTOM_OUTPUT_21072021_keeps <- rbind(ANI_CUSTOM_OUTPUT_21072021_keeps,tmpss)
}



plot_ANI_Percent_shared_species <- ggplot(ANI_CUSTOM_OUTPUT_21072021_keeps,aes(group=ANI_rounded,x=ANI_rounded,shared_spacers))+geom_boxplot(outlier.shape = NA,alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Cas_subtype1)+coord_trans(y="log10")#+geom_point(alpha=0.1,size=0.2)

unique(ANI_CUSTOM_OUTPUT_21072021_keeps$species_ref)
ANI_CUSTOM_OUTPUT_21072021_keeps_cleaned <- ANI_CUSTOM_OUTPUT_21072021_keeps %>% filter(species_ref %in% c("Lactiplantibacillus_plantarum","Lactobacillus_delbrueckii", "Lactobacillus_helveticus","Propionibacterium_freudenreichii","Streptococcus_thermophilus"))

plot_ANI_Percent_shared_species <- ggplot(ANI_CUSTOM_OUTPUT_21072021_keeps,aes(group=ANI_rounded,x=ANI_rounded,y=shared_spacers))+geom_boxplot(outlier.shape = NA,alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Cas_subtype1)+coord_trans(y="log10")+labs(x="ANI (%)",y="shared CRISPR (%)")
plot_ANI_Percent_shared_species
plot_ANI_Percent_shared_species
  png("~/Desktop/Projects/2020_CRISPRscope/Plots//CRISPRshared_ANI_persubtype.png", width = 2100, height = 1500,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)

plot_ANI_Percent_shared_species
 dev.off()

##-----------------only above 99.5

ANI_CUSTOM_OUTPUT_21072021_keeps_highs <- ANI_CUSTOM_OUTPUT_21072021_keeps %>% filter(ANI_rounded>99.4)

ggplot(ANI_CUSTOM_OUTPUT_21072021_keeps_highs,aes(group=ANI_rounded,x=ANI_rounded,shared_spacers))+geom_boxplot(outlier.shape = NA,alpha=0.1,size=0.2)+theme_classic()+facet_wrap(~Cas_subtype1)+coord_trans(y="log10")#+geom_point(alpha=0.1,size=0.2)


```



##CRISPR turn-over rate

```{r}
 ANI_CUSTOM_OUTPUT_21072021 <- read_csv("data_submit/ANI_perArray_sharedSpacers_newer.csv") %>% mutate(shared_spacers=100*perc_shared) %>% filter(shared_spacers!=0) %>% filter(Cas_subtype1==Cas_subtype2) %>% filter(turnover_rate<Inf)

table(ANI_CUSTOM_OUTPUT_21072021$Organism1)
# ANI_CUSTOM_OUTPUT_21072021 <- read_csv("~/Desktop/Projects/2020_CRISPRscope/data/ANI_CUSTOM_OUTPUT_21072021.csv") %>% filter(turnover_rate<Inf) %>% filter(perc_shared!=0)

# ANI_CUSTOM_OUTPUT_21072021$turnover_rate

# ANI_CUSTOM_OUTPUT_21072021$str_typ_1_subtpye <- str_split_fixed(ANI_CUSTOM_OUTPUT_21072021$str_typ_1, fixed("_"), 3)[,3] 
# ANI_CUSTOM_OUTPUT_21072021$str_typ_2_subtpye <- str_split_fixed(ANI_CUSTOM_OUTPUT_21072021$str_typ_2, fixed("_"), 3)[,3] 

ANI_CUSTOM_OUTPUT_21072021$str_typ_1_subtpye <- ANI_CUSTOM_OUTPUT_21072021$Cas_subtype1
ANI_CUSTOM_OUTPUT_21072021$str_typ_2_subtpye <- ANI_CUSTOM_OUTPUT_21072021$Cas_subtype2


table(ANI_CUSTOM_OUTPUT_21072021$str_typ_1_subtpye)
table(ANI_CUSTOM_OUTPUT_21072021$str_typ_2_subtpye)

sum(ANI_CUSTOM_OUTPUT_21072021$str_typ_1_subtpye==ANI_CUSTOM_OUTPUT_21072021$str_typ_2_subtpye)
sum(ANI_CUSTOM_OUTPUT_21072021$str_typ_1_subtpye!=ANI_CUSTOM_OUTPUT_21072021$str_typ_2_subtpye)
ANI_CUSTOM_OUTPUT_21072021[which(ANI_CUSTOM_OUTPUT_21072021$str_typ_1_subtpye!=ANI_CUSTOM_OUTPUT_21072021$str_typ_2_subtpye),]

ANI_CUSTOM_OUTPUT_21072021_cleaned <- ANI_CUSTOM_OUTPUT_21072021 %>% filter(str_typ_1_subtpye==str_typ_2_subtpye)

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned,aes(x=turnover_rate))+geom_density()+theme_classic()+coord_trans(x="log10")#+lims(x=c(0,10000))+ coord_cartesian(clip = 'off')
ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned,aes(x=turnover_rate,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")#+lims(x=c(0,10000))+ coord_cartesian(clip = 'off')

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned,aes(x=turnover_rate,y=perc_shared,fill=str_typ_1_subtpye,color=str_typ_1_subtpye))+geom_point(alpha=0.4)+theme_classic()

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned,aes(x=turnover_rate,y=ANI,fill=str_typ_1_subtpye,color=str_typ_1_subtpye))+geom_point(alpha=0.4)+theme_classic()


ANI_CUSTOM_OUTPUT_21072021_cleaned_subset <- ANI_CUSTOM_OUTPUT_21072021_cleaned %>% filter(ANI>99)

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset,aes(x=turnover_rate,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")#+lims(x=c(0,10000))+ coord_cartesian(clip = 'off')

ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_02 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset %>% filter(turnover_rate<10000) %>% filter(str_typ_1_subtpye!="I-B")



ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_02,aes(x=turnover_rate,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+ coord_cartesian(clip = 'off')+lims(x=c(0,1000))

p1 <- ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_02,aes(x=turnover_rate,fill=str_typ_1_subtpye,color=str_typ_1_subtpye))+geom_density(alpha=0.5)+theme_classic()+coord_trans(x="log10")+ coord_cartesian(clip = 'off')+lims(x=c(0,5000))

ggplotly(p1)

# p <-  ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_02,aes(x=turnover_rate,fill=str_typ_1_subtpye,color=str_typ_1_subtpye))+geom_density(alpha=0.5,adjust=1.5, position="fill")+theme_classic()+coord_trans(x="log10")+ coord_cartesian(clip = 'off')+lims(x=c(0,5000))
# p
# ggplotly(p)



p1 <- ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_02,aes(x=turnover_rate,fill=str_typ_1_subtpye,color=str_typ_1_subtpye))+geom_density(alpha=0.3)+theme_classic()+coord_trans(x="log10")#+ coord_cartesian(clip = 'off')+lims(x=c(0,5000))#+facet_wrap(~str_typ_1_subtpye,scales="free_y")#+ coord_cartesian(clip = 'off')#+lims(x=c(0,10000))+
ggplotly(p1)
p1

turnoversss <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset %>% group_by(str_typ_1_subtpye) %>% summarise(medianss=median(turnover_rate),meanss=mean(turnover_rate),sds=sd(turnover_rate))
turnoversss

##--------final figure

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_02,aes(x=turnover_rate,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(100,250, 500, 1000, 2500,5000,10000))

scale_x_range(breaks=c(2.3,3.2,3.9,4.6), panel = "array size")


#+ coord_cartesian(clip = 'off')#+lims(x=c(0,10000))+ coord_cartesian(clip = 'off')

```

maybe it makes more sense to use the normalized data for this. 
therefore I use the mutation rate of 0.003 and the core genomes size observed in the ANI

```{r}

# core_genome_size <- ANI_CUSTOM_OUTPUT_21072021%>% mutate(core_fraction=bidirectional_fragment_mapping/total_query_fragment) %>% group_by(Cas_subtype2) %>% summarise(meanss=mean(core_fraction),sdss=sd(core_fraction))

# 8.9  1011 per bp per generation

mutation_rate_total=0.003
community_size=10^7
###core-genome normalized mutation rate
ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_03 <- ANI_CUSTOM_OUTPUT_21072021_cleaned %>% mutate(mutation_rate=mutation_rate_total*(bidirectional_fragment_mapping/total_query_fragment)) 


##mean and sd turn over

median(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04$turnover_rate)
mean(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04$turnover_rate)
sd(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04$turnover_rate)

###number of Generations per novel CRISPR
ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_03 %>% mutate(Generations_per_CRISPR=turnover_rate/mutation_rate) %>% mutate(PROBABILITY_per_generation_per_cell=1/Generations_per_CRISPR) 


ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04,aes(x=PROBABILITY_per_generation_per_cell,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=10,hjust=1))

###in community
ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04 %>% mutate(novel_CRISPR_per_generation=community_size/Generations_per_CRISPR) 

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05,aes(x=novel_CRISPR_per_generation,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05,aes(x=novel_CRISPR_per_generation,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~Organism2,scales="free_y")+theme(legend.position = "right",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))

plotx <- ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05,aes(x=novel_CRISPR_per_generation,fill=Organism2))+geom_density(alpha=0.4)+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "right",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))
plotx
ggplotly(plotx)
##--------final figure

ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05 %>% filter(str_typ_1_subtpye!="I-B")
table(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05$str_typ_2_subtpye) %>% sort()

turnoversss <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05 %>% group_by(str_typ_1_subtpye) %>% summarise(medianss=median(novel_CRISPR_per_generation),meanss=mean(novel_CRISPR_per_generation),sds=sd(novel_CRISPR_per_generation)) %>% arrange(.,medianss)
turnoversss


ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06$str_typ_1_subtpye = factor(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06$str_typ_1_subtpye, levels=c(c("I-E","I-C","II-C","I-G","II-A","III-A")))



final_turnover_Rate_plot <- ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06,aes(x=novel_CRISPR_per_generation,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=8,hjust=1),axis.text.y = element_blank(),axis.ticks.y=element_blank())+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))+geom_vline(xintercept = 35,color="darkgrey", linetype="dashed")+labs(y="Density",x="Novel CRISPR spacers per generation in 10^7 cells")

final_turnover_Rate_plot

 # png("~/Desktop/Projects/2020_CRISPRscope/Plots//CRISPR_SPACER_TURNOVER.pdf", width = 1500, height = 1100,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//CRISPR_SPACER_TURNOVER.pdf",width=8,height=5)
# final_turnover_Rate_plot
 # dev.off()

```


use lenski's estimage (8.9  1011 per bp per generation) and account for genome size

```{r}
ANI_CUSTOM_OUTPUT_21072021_cleaned$bidirectional_fragment_mapping/ANI_CUSTOM_OUTPUT_21072021_cleaned$total_query_fragment

mutation_rate_total=8.9 * 10^-11 
community_size=10^7
###core-genome normalized mutation rate
ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_03 <- ANI_CUSTOM_OUTPUT_21072021_cleaned %>% mutate(mutation_rate=(mutation_rate_total*(bidirectional_fragment_mapping*1000))*(bidirectional_fragment_mapping/total_query_fragment)) 

##mean and sd turn over

median(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04$turnover_rate)
mean(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04$turnover_rate)
sd(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04$turnover_rate)

###number of Generations per novel CRISPR
ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_03 %>% mutate(Generations_per_CRISPR=turnover_rate/mutation_rate) %>% mutate(PROBABILITY_per_generation_per_cell=1/Generations_per_CRISPR) 


ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04,aes(x=PROBABILITY_per_generation_per_cell,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=10,hjust=1))

###in community
ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_04 %>% mutate(novel_CRISPR_per_generation=community_size/Generations_per_CRISPR) 

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05,aes(x=novel_CRISPR_per_generation,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))

ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05,aes(x=novel_CRISPR_per_generation,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~Organism2,scales="free_y")+theme(legend.position = "right",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))

plotx <- ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05,aes(x=novel_CRISPR_per_generation,fill=Organism2))+geom_density(alpha=0.4)+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "right",axis.text.x = element_text(angle = 45,size=10,hjust=1))+scale_x_continuous(breaks=c(2,10,50, 100, 250,500))
plotx
ggplotly(plotx)
##--------final figure

ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06 <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05 %>% filter(str_typ_1_subtpye!="I-B")
table(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05$str_typ_2_subtpye) %>% sort()

turnoversss <- ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05 %>% group_by(str_typ_1_subtpye) %>% summarise(medianss=median(novel_CRISPR_per_generation),meanss=mean(novel_CRISPR_per_generation),sds=sd(novel_CRISPR_per_generation)) %>% arrange(.,medianss)
turnoversss


ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_05  %>% summarise(medianss=median(novel_CRISPR_per_generation),meanss=mean(novel_CRISPR_per_generation),sds=sd(novel_CRISPR_per_generation)) 


ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06$str_typ_1_subtpye = factor(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06$str_typ_1_subtpye, levels=c(c("I-E","I-C","II-C","I-G","II-A","III-A")))

ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06$str_typ_1_subtpye = factor(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06$str_typ_1_subtpye, levels=c(c("I-E","I-C","I-G","II-C","II-A","III-A")))

final_turnover_Rate_plot <- ggplot(ANI_CUSTOM_OUTPUT_21072021_cleaned_subset_06,aes(x=novel_CRISPR_per_generation,fill=str_typ_1_subtpye))+geom_density()+theme_classic()+coord_trans(x="log10")+facet_wrap(~str_typ_1_subtpye,scales="free_y")+theme(legend.position = "none",axis.text.x = element_text(angle = 45,size=8,hjust=1),axis.text.y = element_blank(),axis.ticks.y=element_blank())+scale_x_continuous(breaks=c(0.25,0.5,1,2,10,20))+geom_vline(xintercept = 2.8,color="darkgrey", linetype="dashed")+labs(y="Density",x="Novel CRISPR spacers per generation in 10^7 cells")

final_turnover_Rate_plot

 # png("~/Desktop/Projects/2020_CRISPRscope/Plots//CRISPR_SPACER_TURNOVER.pdf", width = 1500, height = 1100,res=300)
#    pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//CRISPR_SPACER_TURNOVER.pdf",width=8,height=5)
# final_turnover_Rate_plot
#  dev.off()

```

# Figure 3

#### Spacer and repeat abundance

```{r}
library(tidyverse)
plot_milread_02 <- read_csv("data_submit/proj_sample_milread_count.csv")
plot_milread_02 <- read_csv("data_submit/proj_sample_milread_count_80.csv") %>% mutate(num_mio_reads_spacer=n_spacer/spacer_per_milread)%>% select(-spacer_per_milread,-repeat_per_milread,) %>% mutate(Spacers=n_spacer_clst80/num_mio_reads_spacer,Repeats=n_repeat_clst80/num_mio_reads_spacer)%>% select(-n_spacer,-n_repeat,-n_spacer_clst80,-n_repeat_clst80,-num_mio_reads_spacer)

metagenomes_sample_speciesRichness_prep <- metagenomes_sample_speciesRichness %>% select(-studyId)

plot_milread_02_ext <- merge(plot_milread_02,metagenomes_sample_speciesRichness_prep,by.x="SRA_ID",by.y="sample_name",all.x=T) %>% as.tibble()


plot_milread_02_ext$group <- plyr::revalue(plot_milread_02_ext$ProjectID,c("Cheese starter cultures"  = "thermophilic",
"Lordan et al. 2019"    = "mesophilic",
"Pasolli et al. 2020"         = "thermophilic",
"Swiss raclette"        = "mesophilic",
"Walsh et al. 2020"     = "mesophilic"))


plot_milread_0_long <- gather(plot_milread_02_ext, type, per_mil, c("Spacers","Repeats"), factor_key=TRUE,na.rm = TRUE) %>% select(-ProjectID)



##------------
##stats
##------------
 plot_milread_0_long %>% group_by(type) %>% summarise(means=mean(per_mil),sdss=sd(per_mil))


###------------------------
#include mnch et al. 
###------------------------

CRISPR_abundance_munch_et_al <- read_delim("data_submit/CRISPR_abundance_munch_et_al.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

CRISPR_abundance_munch_et_al_long <- gather(CRISPR_abundance_munch_et_al, type, per_mil, c("Spacers","Repeats"), factor_key=TRUE,na.rm = TRUE) %>% select("SRA_ID" , "group" , "species_richness", "type", "per_mil"  )

plot_milread_0_long_complet <- rbind(plot_milread_0_long,CRISPR_abundance_munch_et_al_long) %>% mutate(grouping="normalized by\nsequecing depth")

plot_milread_0_long_complet$group = factor(plot_milread_0_long_complet$group, levels=rev(c("mesophilic","thermophilic","mnch et al, 2021")))


###------------------------
#new include mnch et al. 
###------------------------

CRISPR_abundance_munch_et_al_species <- read_csv("data_submit/CRISPR data - Figure 2 - Panel C (x).csv")
CRISPR_abundance_munch_et_al_spacers <- read_csv("data_submit/CRISPR data - Figure 2 - Panel C (y) extended.csv")
colnames(plot_milread_0_long)
CRISPR_abundance_munch_et_al_long <- merge(CRISPR_abundance_munch_et_al_spacers,CRISPR_abundance_munch_et_al_species,by.x="sample",by.y="Species") %>% as_tibble() %>% mutate(group="mnch et al, 2021")%>% mutate(type="Spacers") %>% select(sample,`Number (> 1%RA)`,group,type,spacers_80) 

colnames(CRISPR_abundance_munch_et_al_long) <- c("SRA_ID","species_richness","group","type"  , "per_mil" )

# CRISPR_abundance_munch_et_al_long <- gather(CRISPR_abundance_munch_et_al, type, per_mil, c("Spacers","Repeats"), factor_key=TRUE,na.rm = TRUE) %>% select("SRA_ID" , "group" , "species_richness", "type", "per_mil"  )

plot_milread_0_long_complet_01 <- rbind(plot_milread_0_long,CRISPR_abundance_munch_et_al_long) %>% mutate(grouping="normalized by\nsequecing depth")

plot_milread_0_long_complet_02 <- rbind(plot_milread_0_long,CRISPR_abundance_munch_et_al_long) %>% mutate(grouping="normalized by\nrichness") %>% mutate(per_mil=per_mil/species_richness) %>% filter(!is.na(per_mil))
plot_milread_0_long_complet <- rbind(plot_milread_0_long_complet_01,plot_milread_0_long_complet_02) %>% filter(type!="Repeats")  %>% as_tibble()

plot_milread_0_long_complet$group = factor(plot_milread_0_long_complet$group, levels=rev(c("mesophilic","thermophilic","mnch et al, 2021")))

plot_milread_0_long_complet$grouping = factor(plot_milread_0_long_complet$grouping, levels=(c("normalized by\nsequecing depth","normalized by\nrichness")))

# colorssss <- rev(c("blue","red","grey77"))

# colorssss <- rev(c("#B1DDF1","#9F87AF","grey77"))
colorssss <- rev(c("#6DAEDB","#FFB4A2","grey77"))
# colorssss <- rev(c("#A9B3CE","#9E788F","grey77"))
# colorssss <- rev(c("#465775","#EF6F6C","grey77"))


###------------------------
#plot_normalised per million
###------------------------

# plot_perMio <- ggplot(plot_milread_0_long_complet,aes(x=group,y=per_mil,fill=group,color=group))+geom_violin(alpha=0.6,adjust=0.8)+theme_classic()+labs(x="",y="Observed number per million")+ggforce::geom_sina(alpha=0.8)+ coord_flip()+theme(legend.position = "none")+facet_grid(grouping~type,scales = "free")+geom_boxplot(color="grey44",alpha=0,width=.25)+scale_fill_manual(values =colorssss )+scale_color_manual(values =colorssss )
# plot_perMio

###------------------------
#add richness
###------------------------

# plot_milread_0_long_RICHNESS <- plot_milread_0_long_complet %>% filter(!is.na(species_richness))%>% mutate(per_mil=per_mil/species_richness) %>% mutate(grouping="normalized by\nRichness")

plot_milread_0_long_complet_1 <- plot_milread_0_long_complet %>% filter(grouping=="normalized by\nsequecing depth")

plot_richness_01 <- ggplot(plot_milread_0_long_complet_1,aes(x=group,y=per_mil,fill=group,color=group))+geom_violin(alpha=0.6,adjust=0.8)+theme_classic()+labs(x="",y="Normalised number per million")+ggforce::geom_sina(alpha=0.8)+ coord_flip()+theme(legend.position = "none")+facet_grid(grouping~.,scales = "free_x")+geom_boxplot(color="grey44",alpha=0,width=.25)+scale_fill_manual(values =colorssss )+scale_color_manual(values =colorssss )
plot_richness_01

plot_milread_0_long_complet_2 <- plot_milread_0_long_complet %>% filter(grouping!="normalized by\nsequecing depth")


plot_richness_02 <- ggplot(plot_milread_0_long_complet_2,aes(x=group,y=per_mil,fill=group,color=group))+geom_violin(alpha=0.6,adjust=0.8)+theme_classic()+labs(x="",y="Normalised number per million")+ggforce::geom_sina(alpha=0.8)+ coord_flip()+theme(legend.position = "none")+facet_grid(grouping~.,scales = "free_x")+geom_boxplot(color="grey44",alpha=0,width=.25)+scale_fill_manual(values =colorssss )+scale_color_manual(values =colorssss )
plot_richness_02

###------------------------
#merge plots
###------------------------

# plot_perMio+plot_richness+  plot_layout(heights = c(1, 1))
 # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/counts_spacer_repeat_samples_02.pdf",width=4.5,height=5)
plot_richness_01+plot_richness_02+  plot_layout(heights = c(1, 1))

# dev.off()
##--------------
#significance
##--------------

 stat.test<- plot_milread_0_long_complet %>%
  group_by(grouping,type) %>%
  t_test(per_mil ~ group)  %>%adjust_pvalue(method = "bonferroni") %>%
  add_significance()
stat.test 


for (groupingzzz in as.character(unique(plot_milread_0_long_complet$grouping))) {
  
  for (groupss1 in  as.character(unique(plot_milread_0_long_complet$group))) {
    
          tmp <- plot_milread_0_long_complet %>% filter(grouping==groupingzzz) %>% filter(group != groupss1)

    
          groupzz1 <- unique(tmp$group)[1]
          groupzz2 <- unique(tmp$group)[2]
          
          
          tmp$group <- droplevels(tmp$group)
    # for (groupss2 in  as.character(unique(plot_milread_0_long_complet$group))) {
    
                # tmp1 <- tmp %>% filter(grouping==groupingzzz) %>% filter(group != groupss2)

      # unique(tmp$group)
      
      tmp3 <- wilcox.test(per_mil~group, data=tmp, paired=FALSE)
      # tmp3$p.value
       tmp4 <- data.frame(grouping=groupingzzz,group1=groupzz1,group2=groupzz2,pvalue=tmp3$p.value)
      
       wilcoxen_grouping <- rbind(wilcoxen_grouping, tmp4)
       

  }
  

}

```




## rarefaction spacer

Here, I look if the metagenomic spacers are presence. 

```{r}

cluster_by_sample_spread_01 <- read_csv("data_submit/cluster_by_sample_spread_01.csv") #,   col_types = cols(`1944` = col_character())
# str(cluster_by_sample_spread_01)

cluster_by_sample_spread_01 %>% mutate()

# colSums(cluster_by_sample_spread_01$`1944`)
# sum(ifelse(cluster_by_sample_spread_01$`1944`)

ifelse(cluster_by_sample_spread_01$`1944`)

cluster_by_sample_spread_01$`1944` <- as.numeric(ifelse(cluster_by_sample_spread_01$`1944`=="0","0","1"))


for (variabless in 2:ncol(cluster_by_sample_spread_01)) {
  
cluster_by_sample_spread_01[,variabless] <- as.numeric(ifelse(cluster_by_sample_spread_01[,variabless]=="0","0","1"))

}

colSums(cluster_by_sample_spread_01[,-1]) %>% hist()
colnames(cluster_by_sample_spread_01)
cluster_by_sample_spread_preped <- cluster_by_sample_spread_01 %>% remove_rownames() %>% column_to_rownames(., var = "cluster_spacer_identity") %>% t()

#%>% mutate(cluster_spacer_identity=paste0("cluster",cluster_spacer_identity))

# cluster_by_sample_spread_preped <- cluster_by_sample_spread_01 %>% remove_rownames()%>% column_to_rownames(., var = "cluster_spacer_identity")

colSums(cluster_by_sample_spread_preped) %>% hist()

cluster_by_sample_spread_preped <- cluster_by_sample_spread_preped[which(!is.na(rowSums(cluster_by_sample_spread_preped))),]
cluster_by_sample_spread_spacer <- cluster_by_sample_spread_preped

dim(cluster_by_sample_spread_spacer)
###------------------------------------
# rarefaction plot
##-----------------------------

# gene_presence_absence_roary_preped <- gene_presence_absence_roary[,-c(2:14)] %>% remove_rownames()%>% column_to_rownames(., var = "Gene")
# hist(colSums(cluster_by_sample_spread_preped),xlim = c(0,10),breaks = 1:100)

rar.tbl <- rarefaction(cluster_by_sample_spread_preped, n.perm = 200)
# data_genome_all <- rar.tbl %>%   gather(key = "Permutation", value = "Clusters", -Genome)  #%>% filter(!is.na(Clusters))
data_genome_all <- rar.tbl %>%   gather(key = "Permutation", value = "Clusters", -Genome) %>% add_column(type="spacers")
data_genome_spacers <- data_genome_all

# rarefy_test <- rarefy(cluster_by_sample_spread_preped,sample=200)

# rarefy_test <- rarefy(cluster_by_sample_spread_preped,min(rowSums(cluster_by_sample_spread_preped)))

max(data_genome_all$Clusters)

Rarefaction_all <- ggplot(data_genome_all,aes(x = Genome, y = Clusters))+
     geom_smooth(method="nls",formula=y~s*x/(F+x),method.args=list(start = c( F = 10,s=1000 )), se=FALSE)+theme_classic()+
    labs(x="Metagenomic samples",y="CRISPR spacers")
Rarefaction_all

   pdf("~/Desktop/Projects/2020_CRISPRscope/Plots/rarefactions_spacers.pdf",width=3.5,height=3.5)

Rarefaction_all

dev.off()


#------------------------------------PCOA

## Principal coordinate analysis and simple ordination plot
mite.D <- vegdist(cluster_by_sample_spread_preped, "bray")
res <- pcoa(mite.D)
res$values
biplot(res)

## Project unstandardized and standardized species on the PCoA ordination plot
mite.log.st = apply(mite.D, 2, scale, center=TRUE, scale=TRUE)

par(mfrow=c(1,2))
biplot(res, mite.log)
biplot(res, mite.log.st)


par(mfrow=c(1,1))

##test

df_pca <- prcomp(cluster_by_sample_spread_preped)

plot(df_pca$x[,1], df_pca$x[,2])





df_pca <- prcomp(defense_systems_complete_wide_preped)

plot(df_pca$x[,1], df_pca$x[,2])

  


metabolo.sum <- summary(df_pca)
metabolo.sum
metabolo.sum <- metabolo.sum$importance %>% as.data.frame()

ploting_data <- df_pca$x %>% as.data.frame()
ploting_data$genome <- rownames(ploting_data)
ploting_data_final <-  merge(ploting_data,defense_systems_species,by="genome")

table(ploting_data_final$species)

pca0 <- ggbiplot(df_pca,var.axes=FALSE,obs.scale = 1)+theme_classic()+labs(title="grouping by biological replicates")+theme(legend.position = "none")
pca0

pca1 <- ggbiplot(df_pca,var.axes=FALSE,groups=ploting_data_final$species,obs.scale = 1,alpha=0.3,ellipse=TRUE)+theme_classic()+theme(legend.position = "right")
pca1

species_colors <- read_delim("data_submit/species_colors.txt", "\t", escape_double = FALSE, trim_ws = TRUE)

```


# Figure 4

## Protspacer vs. spacers



```{r}
##-----------------------------------------------
##------------------------import data
##-----------------------------------------------

DR_SP_DR_info <- read_csv("data_submit/DR_SP_DR_info.csv", col_types = cols(sample = col_character())) %>% filter(sample %in% unique(X20210424_metaphlan$sample_name))  #%>%  select(-readNumber,-cluster_repeat_80,-cluster_spacer_identity,-cluster_repeat_identity,-cluster_repeat_identity,-cluster_spacer_80)
table(DR_SP_DR_info$grouping)
 


DR_SP_DR_info %>% nrow()
DR_SP_DR_info %>% filter(spacer_cov<3) %>% nrow()
DR_SP_DR_info %>% filter(protospacer_cov<3) %>% nrow()
DR_SP_DR_info %>% filter(spacer_cov>3) %>% filter(protospacer_cov>3)%>% nrow()  


 100*((DR_SP_DR_info$cluster_spacer_80 %>% unique() %>% length())/ 8226 )
 
 
 table(DR_SP_DR_info$sample)
  
  DR_SP_DR_info$grouping <- ifelse(DR_SP_DR_info$spacer_cov<3,"kein spacer",ifelse(DR_SP_DR_info$protospacer_cov>3,"mit proto","ohne proto"))

  
  DR_SP_DR_info %>% filter(grouping=="mit proto") %>% select(sample,cluster_spacer_80) %>% unique() %>% group_by(sample) %>% summarise(countss=n())
    DR_SP_DR_info %>% filter(grouping=="ohne proto") %>% select(sample,cluster_spacer_80) %>% unique() %>% group_by(sample) %>% summarise(countss=n())

  
 DR_SP_DR_info %>% filter(spacer_cov>3) %>% nrow()
 
 
  

CRISPR_spacer_coverage_filtered <- DR_SP_DR_info %>% filter(spacer_cov>3) %>% filter(protospacer_cov>3) %>%  select(-readNumber,-cluster_repeat_80,-cluster_spacer_identity,-cluster_repeat_identity,-cluster_repeat_identity,-cluster_spacer_80)

CRISPR_spacer_coverage_filtered <- CRISPR_spacer_coverage %>% filter(spacer_cov>3) %>% filter(protospacer_cov>3)


# CRISPR_spacer_coverage_filtered,DR_SP_DR_info,by.x="spacer_name",by.y=""

my.formula <- y ~ x

plot <- ggplot(CRISPR_spacer_coverage_filtered,aes(x=spacer_CPM,y=protospacer_CPM))+geom_point(alpha=0.2)+theme_classic()+
# coord_trans(y="log2",x="log2")+
    scale_x_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  scale_y_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  # lims(x=c(0,100000),y=c(0,100000))+
  labs(x="Spacer (cpm)",y="Protospacers  (cpm)")+
geom_smooth(method="lm",se = FALSE,color="blue",linetype = "dashed")+
  stat_poly_eq(formula = my.formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               parse = TRUE,)

# + geom_abline(intercept = 1060, slope = 0.508,color="red")
plot

#-------------------important subtypes
###suptypes 

CRISPR_spacer_coverage_filtered_subs <- CRISPR_spacer_coverage_filtered %>% filter(Subtype %in% c("I-E","I-C","II-C","I-G","II-A","III-A"))

  
plot2 <- ggplot(CRISPR_spacer_coverage_filtered_subs,aes(x=spacer_CPM,y=protospacer_CPM,color=Subtype))+geom_point(alpha=0.2)+theme_classic()+
# coord_trans(y="log2",x="log2")+
    scale_x_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  scale_y_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  # lims(x=c(0,100000),y=c(0,100000))+
  labs(x="Spacer (cpm)",y="Protospacers  (cpm)")+
geom_smooth(method="lm",se = FALSE)

plot2

plot+plot2
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//protospacer_vs_spacer.pdf",width=9,height=4)
# plot+plot2
 # dev.off()

 ##------------------
 #de-seq plot
 ##------------------
 
 CRISPR_spacer_coverage_filtered_de_seq <- CRISPR_spacer_coverage_filtered %>% mutate(deseq_count= protospacer_CPM/spacer_CPM)
 
 CRISPR_spacer_coverage_filtered_de_seq %>% summarise(means=mean(deseq_count),medians=median(deseq_count))
 
 
 plot3 <- ggplot(CRISPR_spacer_coverage_filtered_de_seq,aes(x=spacer_CPM,y=deseq_count))+geom_point(alpha=0.2)+theme_classic()+
# coord_trans(y="log2",x="log2")+
    scale_x_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  scale_y_log10()+
  # lims(x=c(0,100000),y=c(0,100000))+
  labs(x="Spacer (cpm)",y="Protospacers vs. spacer  (cpm)")+
geom_smooth(method="lm",se = FALSE,color="blue",linetype = "dashed")#+
  # stat_poly_eq(formula = my.formula,
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
  #              parse = TRUE,)

# + geom_abline(intercept = 1060, slope = 0.508,color="red")
plot3

 # png("~/Desktop/Projects/2020_CRISPRscope/Plots/viral_microbe_ratio.png", width = 1200, height = 1200,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)
# plot3
 # dev.off()
 
#-------------------important subtypes
###suptypes 

CRISPR_spacer_coverage_filtered_de_seq_subs <- CRISPR_spacer_coverage_filtered_de_seq %>% filter(Subtype %in% c("I-E","I-C","II-C","I-G","II-A","III-A"))

  
plot4 <- ggplot(CRISPR_spacer_coverage_filtered_de_seq_subs,aes(x=spacer_CPM,y=deseq_count,color=Subtype))+geom_point(alpha=0.2)+theme_classic()+
# coord_trans(y="log2",x="log2")+
    scale_x_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  scale_y_log10()+
  # lims(x=c(0,100000),y=c(0,100000))+
  labs(x="Spacer (cpm)",y="Protospacers  (cpm)")+
geom_smooth(method="lm",se = FALSE)

plot4

plot3+plot4
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//protospacer_vs_spacer_02.pdf",width=9,height=4)
# plot3+plot4
 # dev.off()
 
 
 
 ###-------------------------------
 #abundacne of protospacer only
 
 CRISPR_spacer_coverage %>% filter(spacer_cov>3 | protospacer_cov>3) %>% nrow()
 
 CRISPR_spacer_coverage_filtered_proto <- CRISPR_spacer_coverage %>% filter(spacer_cov<3) %>% filter(protospacer_cov>3) %>% add_column(groupingss="protoONLY")
 CRISPR_spacer_coverage_filtered_both <- CRISPR_spacer_coverage %>% filter(spacer_cov>3) %>% filter(protospacer_cov>3) %>% add_column(groupingss="both")

 combined <- rbind(CRISPR_spacer_coverage_filtered_proto,CRISPR_spacer_coverage_filtered_both)
 
 median(CRISPR_spacer_coverage_filtered_proto$protospacer_CPM)
 mean(CRISPR_spacer_coverage_filtered_proto$protospacer_CPM)
 
 plot5 <- ggplot(combined,aes(x=groupingss,y=protospacer_CPM))+geom_boxplot()+theme_classic()+
# coord_trans(y="log2",x="log2")+
    #scale_x_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  scale_y_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))+
  # lims(x=c(0,100000),y=c(0,100000))+
   # coord_trans(y="log10")+
  labs(y="Protospacers  (cpm)")
# geom_smooth(method="lm",se = FALSE,color="blue",linetype = "dashed")+
  # stat_poly_eq(formula = my.formula,
               # aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               # parse = TRUE,)

# + geom_abline(intercept = 1060, slope = 0.508,color="red")
plot5

 ggplotly(plot5)
 
 ##--------------------------------------
 #abundance of spacers within sampels
 
 
CRISPR_spacer_coverage_filtered <-  CRISPR_spacer_coverage %>% as_tibble() %>% filter(spacer_CPM>0 || protospacer_CPM>0) %>% filter(sample %in% unique(X20210424_metaphlan$sample_name))

 
 coiunts_sp_dr_sp_sum <- CRISPR_spacer_coverage %>% filter(sample %in% unique(X20210424_metaphlan$sample_name))%>% group_by(sample) %>% summarise(counts=n()) 

range(coiunts_sp_dr_sp_sum$counts)
median(coiunts_sp_dr_sp_sum$counts)
mean(coiunts_sp_dr_sp_sum$counts)
sd(coiunts_sp_dr_sp_sum$counts)

CRISPR_spacer_coverage$sample %>% unique()
coiunts_sp_dr_sp_sum %>% filter(counts < 2)

unique(X20210424_metaphlan$sample_name)



ggplot(CRISPR_spacer_coverage_filtered,aes(fill=sample,x=sample,y=spacer_CPM))+geom_boxplot()+geom_jitter()+theme_classic()+theme(legend.position = "none")+scale_y_log10(limit=c(1e+01,1e+05),breaks=c(1e+01,1e+03,1e+05))


###-normalise

CRISPR_spacer_coverage_filtered_max <- CRISPR_spacer_coverage_filtered %>% group_by(sample) %>% summarise(normcount=max(spacer_CPM)) %>% left_join(.,CRISPR_spacer_coverage_filtered,) %>% mutate(norm_spacers=(spacer_CPM)/normcount) %>% filter(spacer_CPM>0) 


spacer_abundance_plot <- ggplot(CRISPR_spacer_coverage_filtered_max,aes(fill=sample,x=sample,y=norm_spacers))+geom_boxplot()+theme_classic()+theme(legend.position = "none",axis.text.x = element_blank())+labs(y="normalized spacer abundance",x="samples")
spacer_abundance_plot
  # png("~/Desktop/Projects/2020_CRISPRscope/Plots/spacer_abundance_plot.png", width = 1800, height = 1200,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)
# spacer_abundance_plot
 # dev.off()
# +geom_jitter(alpha=0.1
 
# library(hrbrthemes)
 

spacer_abundance_plot_ridge <- ggplot(CRISPR_spacer_coverage_filtered_max,aes(y=sample,x=norm_spacers,fill=..x..))+geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01)  +scale_fill_viridis() +theme_classic()+theme(legend.position = "none",axis.text.y = element_blank())+labs(x="normalized spacer abundance",y="Metagenomes",title = "Metagenomes contain between 41 and 1961 spacers")
spacer_abundance_plot_ridge

 ##---------------------------subset for only spacers from genomes
DR_SP_DR_info <- read_csv("data_submit/DR_SP_DR_info.csv", col_types = cols(sample = col_character())) %>% filter(sample %in% unique(X20210424_metaphlan$sample_name))  #%>%  select(-readNumber,-cluster_repeat_80,-cluster_spacer_identity,-cluster_repeat_identity,-cluster_repeat_identity,-cluster_spacer_80)

DR_SP_DR_info[is.na(DR_SP_DR_info$cluster_spacer_identity),]
 
 DR_SP_DR_info_spacer_origin <-  DR_SP_DR_info %>% select(spacer_name,cluster_spacer_identity) %>% unique()

 CRISPRscope_meta <- read_csv("data_submit/20210125_CRISPRscope_meta.csv",col_types = cols(SRA_ID = col_character())) %>% mutate(spacer_name=paste(ProjectID,SRA_ID,GID,SPID,Coverage,sep="_")) %>% select(spacer_name,cluster_spacer,identity_spacer_cluster)
 
sum(is.na(CRISPRscope_meta$identity_spacer_cluster))
 sum(is.na(CRISPRscope_meta$cluster_spacer))

 CRISPR_spacer_coverage_filtered_max_extended <- CRISPR_spacer_coverage_filtered_max %>% left_join(.,DR_SP_DR_info_spacer_origin)
# sum(is.na(CRISPR_spacer_coverage_filtered_max_extended$cluster_spacer_identity))
 OrganismStrainClustersp <- read_csv("data_submit/OrganismStrainClustersp.csv")

 CRISPR_spacer_coverage_filtered_max_extended_strains <- CRISPR_spacer_coverage_filtered_max_extended %>% filter(cluster_spacer_identity %in% unique(OrganismStrainClustersp$cluster_spacer_identity))
 
 spacer_abundance_plot_ridge <- ggplot(CRISPR_spacer_coverage_filtered_max_extended_strains,aes(y=sample,x=norm_spacers,fill=..x..))+geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01)  +scale_fill_viridis() +theme_classic()+theme(legend.position = "none",axis.text.y = element_blank())+labs(x="normalized spacer abundance",y="Metagenomes",title = "Metagenomes contain between 41 and 1961 spacers")
spacer_abundance_plot_ridge
meta_sp_protosp_coverage_hits <- read.csv("data_submit/meta_sp_protosp_coverage_hits.csv")  %>% filter(spacer_CPM > 0 | protospacer_CPM > 0 )

meta_sp_protosp_coverage_hits %>%  select(spacer_name,cluster_spacer_identity) %>% unique()%>% filter(!is.na(cluster_spacer_identity))

 
 ##---------------------------spacer shared amongst metaG
 
 CRISPR_spacer_coverage_filtered_counts <- CRISPR_spacer_coverage_filtered %>% filter(spacer_CPM>0) %>% group_by(spacer_name) %>% summarise(Number_of_samples=n())
 table(CRISPR_spacer_coverage_filtered_counts$Number_of_samples)
 
 length(CRISPR_spacer_coverage_filtered_counts$Number_of_samples) 

(6293+397)/7313

 
 spacer_shared_metaG <- ggplot(CRISPR_spacer_coverage_filtered_counts,aes(x=Number_of_samples))+geom_bar()+theme_classic()+labs(y="Spacers",x="Number of metagenomes shared")
spacer_shared_metaG
 
```

# Figure 5


# Blast spacers

Here, I blast the spacers against

do blasting against:

1. my viral DB
2. IMG viral DB
3. REFSEQ db
4. protein DB

```{bash}

mappingFile=/data/projects/p539_crisprscope/CRISPRblast/cluster1_spacers_reference_CRISPRscope.fasta
BaseLocation=/data/projects/p539_crisprscope/CRISPRblast/
mkdir -p ${BaseLocation}/blast_02

  blastn -num_threads 8 -max_hsps 1 -max_target_seqs 1 -task megablast -show_gis \
  -query ${mappingFile} \
  -outfmt "6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle slen sgi sacc sblastnames staxids sscinames" \
  -db /archiv/Phage_DB/BLAST/20190519/Vcontact/Database/alltogether/data/merged_all_phages \
  -out ${BaseLocation}/blast_02/CRISPRblast_ViralDB.txt \
  -evalue 0.01 -word_size 16
  

##--------------------------blast to old viral DB
#this is the RefSeq database

  blastn -num_threads 35 -max_hsps 1 -max_target_seqs 1 -task megablast -show_gis \
  -query ${mappingFile} \
  -outfmt "6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle slen sgi sacc sblastnames staxids sscinames" \
  -db /archiv/BLAST/blast_db/20190204/viral_nucleotide/viral \
  -out ${BaseLocation}/blast_02/CRISPRblast_ViralDB_old.txt \
  -evalue 0.01 -word_size 16
  
  
##--------------------------blast unmapped to nucleotide DB


blastn -num_threads 37 -max_hsps 1 -max_target_seqs 1 -task megablast -show_gis \
  -query ${mappingFile} \
  -outfmt "6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle slen" \
  -db /archiv/BLAST/blast_db/20190204/nucleotide/prokaryotes \
  -out  ${BaseLocation}/blast_02/CRISPRblast_nucleotideDB.txt \
  -evalue 0.01 -word_size 16

wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB.txt 
wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB_old.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_nucleotideDB.txt

##--------------------------img/vr
cd /home/vincent/anaconda3/envs/PROKKA/db/hmm
#makeblastdb -dbtype nucl -in IMGVR_all_nucleotides.fna

blastn -num_threads 37 -max_hsps 1 -max_target_seqs 1 -task megablast -show_gis \
  -query ${mappingFile} \
  -outfmt "6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle slen" \
  -db /work/Databases/20210128_img_vr/IMGVR_all_nucleotides.fna \
  -out  ${BaseLocation}/blast_02/CRISPRblast_IMGVR.txt \
  -evalue 0.01 -word_size 16

wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB.txt 
wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB_old.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_nucleotideDB.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_IMGVR.txt

##--------------------------nt
#cd /home/vincent/anaconda3/envs/PROKKA/db/hmm
#makeblastdb -dbtype nucl -in IMGVR_all_nucleotides.fna
conda activate blast

blastn -num_threads 37 -max_hsps 1 -max_target_seqs 1 -task megablast -show_gis \
  -query ${mappingFile} \
  -outfmt "6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle slen staxid ssciname sskingdom staxids" \
  -db /work/Databases/20210708_nt/nt \
  -out  ${BaseLocation}/blast_02/CRISPRblast_nt.txt \
  -evalue 0.01 -word_size 16



wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB.txt 
wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB_old.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_nucleotideDB.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_IMGVR.txt

```

```{r}

CRISPRblast_ViralDB <- read_delim("data_submit/CRISPRblast_ViralDB.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE, col_types = cols(X1 = col_character())) %>% select(c(X1,X13)) %>% rename(., "spacer"="X1")%>% rename(., "ViralDB"="X13")

CRISPRblast_ViralDB_old <- read_delim("data_submit/CRISPRblast_ViralDB_old.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE, col_types = cols(X1 = col_character())) %>% select(c(X1,X13)) %>% rename(., "spacer"="X1")%>% rename(., "ViralDB_old"="X13")

CRISPRblast_nucleotideDB <- read_delim("data_submit/CRISPRblast_nucleotideDB.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE, col_types = cols(X1 = col_character()))%>% select(c(X1,X13)) %>% rename(., "spacer"="X1")%>% rename(., "nucleotideDB"="X13")

CRISPRblast_IMGVR <- read_delim("data_submit/CRISPRblast_IMGVR.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE, col_types = cols(X1 = col_character()))%>% select(c(X1,X13)) %>% rename(., "spacer"="X1")%>% rename(., "IMGVR"="X13")


cluster_reference_CRISPRscope <- read_csv("data_submit/cluster_reference_CRISPRscope.txt",  col_names = FALSE, col_types = cols(X1 = col_character()))%>% rename(., "spacer"="X1")

# multiFull <- merge(  cluster_reference_CRISPRscope,  CRISPRblast_ViralDB_old, all = TRUE),by="spacer",  CRISPRblast_nucleotideDB, all = TRUE),by="spacer",  CRISPRblast_ViralDB, all = TRUE),by="spacer",
  # CRISPRblast_IMGVR, all = TRUE)
i=1
i=10
for (i in 1:nrow(multiFull)) {
  ifelse(!is.na(multiFull[,"CRISPRblast_ViralDB_old"])|!is.na(multiFull["CRISPRblast_ViralDB",i])|!is.na(multiFull["CRISPRblast_IMGVR",i])

}
multiFull[1,"ViralDB_old"]

 multiFull$final <-  ifelse(!is.na(multiFull[,"ViralDB_old"])|!is.na(multiFull[,"ViralDB"])|!is.na(multiFull[,"IMGVR"]),"phage",ifelse(!is.na(multiFull[,"nucleotideDB"]),"bacteria","No-hit"))
  table(multiFull$final)
  
  
  # write.csv(multiFull,file="~/Desktop/Projects/2020_CRISPRscope/05_BCGtree/CRISPRblast/blast//final_spacerHits.txt")

```

## Identify CRISPR in blast

Here, we want to see if any of the BLAST hits are actually still CRISPRs

```{bash}

awk -F "," '{if ($7=="Bacterial")print $2}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv |head

awk -F "," '{if ($7=="Bacterial")print $2}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv |cut -d '|' -f 2 |head



wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv

grep "nt,Bacterial" -c /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv 
grep "nt,Viral" -c /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv 

grep "nt,Bacterial"  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv |cut -d ',' -f 2 |head

awk -F "," '{if ($7=="Bacterial")print $2}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv |wc -l
awk -F "," '{if ($7=="Bacterial")print $2}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv |cut -d '|' -f 2|sort|uniq|wc -l

###------------------save entrez gi numbers for batch search

grep "nt,Bacterial"  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories.csv |cut -d ',' -f 2  |cut -d '|' -f  2 |sort |uniq > /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories_gi_num.csv

###------------------NEW FILE: save entrez gi numbers for batch search

wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv

awk -F "," '{if($2=="Bacterial")print $3}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv |cut -d '|' -f 2 |wc -l 

awk -F "," '{if($2=="Bacterial")print $3}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv |cut -d '|' -f 2 |sort|uniq |wc -l 

awk -F "," '{if($2=="Bacterial")print $3}' /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv |cut -d '|' -f 2 |sort|uniq > /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories_gi_num.csv

#grep "1567597754" /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories_gi_num.csv
#grep "sseqid" /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories_gi_num.csv

####----------------------summary from all gi results

#/home/vincent/Desktop/Projects/2020_CRISPRscope/data/nuccore_result.txt
rm /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
for idsss in $(cat /home/vincent/Desktop/Projects/2020_CRISPRscope/data/Blast_nt_results_categories_gi_num.csv)
do

match=$(grep "GI:${idsss}$" -B 2 /home/vincent/Desktop/Projects/2020_CRISPRscope/data/nuccore_result.txt |head -1)
echo -e ${idsss}"\t"${match} >> /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
done

###---------summary
wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "plasmid" -i -c /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "crispr" -i -c  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "phage" -i -c  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "virus" -i -c  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "homo" -i -c  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "mus" -i -c  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt
grep "zebra" -i -c  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt


grep "plasmid" -i -v /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt|grep "crispr" -v -i|grep "phage" -i -v|grep "virus" -i -v |wc -l


grep "plasmid" -i -v /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt|grep "crispr" -v -i|grep "phage" -i -v|grep "virus" -i -v  |cut -f 1 > /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names_genomes.txt


##-----------------------identify how many CRISPR (originally bacterial targeting) spacers are from CRISPR or targeting plasmids

rm /home/vincent/Desktop/Projects/2020_CRISPRscope/data/plasmid_hits.txt
for idss in $(grep "plasmid" -i /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt |cut -f 1)
do

counts=$(grep "gi|${idss}|" -c /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv)
echo -e ${idss}"\t"${counts}
grep "gi|${idss}|"  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv >> /home/vincent/Desktop/Projects/2020_CRISPRscope/data/plasmid_hits.txt
done
wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/plasmid_hits.txt


##----
rm /home/vincent/Desktop/Projects/2020_CRISPRscope/data/CRISPR_hits.txt
for idss in $(grep "CRISPR" -i /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names.txt |cut -f 1)
do

counts=$(grep "gi|${idss}|" -c /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv)
echo -e ${idss}"\t"${counts}

grep "gi|${idss}|"  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv >> /home/vincent/Desktop/Projects/2020_CRISPRscope/data/CRISPR_hits.txt
done
wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/CRISPR_hits.txt

##---------------------download non-CRISPR/non-plasmid genomes
curl 'ftp://ftp.ncbi.nlm.nih.gov/genomes/refseq/bacteria/assembly_summary.txt'  > /home/vincent/Desktop/Projects/2020_CRISPRscope/download.file
grep "AJ004687.2" /home/vincent/Desktop/Projects/2020_CRISPRscope/download.file


###----------------------------------------
#prepare non-crispr and non-plasmid entrez file
###----------------------------------------
wc -l  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names_genomes.txt


rm /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS.txt
for idss in $(cat /home/vincent/Desktop/Projects/2020_CRISPRscope/data/gi_sequence_names_genomes.txt)
do

grep "gi|${idss}|"  /home/vincent/Desktop/Projects/2020_CRISPRscope/data/CRISPRblast_nt_vincent.txt >> /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS.txt
done
wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS.txt

##------------subset to only bacterial
rm /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS_curated.txt
for clustersss in $(cut -d ',' -f 1 /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv|sed '1d')
do

grep "^${clustersss}$(printf '\t')" /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS.txt >> /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS_curated.txt


done

wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS_curated.txt
wc -l /home/vincent/Desktop/Projects/2020_CRISPRscope/data/BLAST_Bacterial_hits_target_id.csv

```


```{bash}


scp /home/vincent/Desktop/Projects/2020_CRISPRscope/data/non_crispr_or_plasmid_blastHITS_curated.txt vincent@130.223.51.66:/data/projects/p539_crisprscope/CRISPRblast//blast_02/
```


download from entrez

```{bash}
##blast ouput

less /home/vincent/Desktop/Projects/2020_CRISPRscope/data/CRISPRblast_nt_vincent.txt

##----------test entrez
mkdir -p /data/projects/p539_crisprscope/entrez_download
cd /data/projects/p539_crisprscope/entrez_download

queryiesss=1015427617
startsss=10
endsss=2000

#esearch -db nucleotide -query "${queryiesss}" |  efetch -format fasta -seq_start ${startsss} -seq_stop ${endsss}

#esearch -db nucleotide -query "${queryiesss}" |  efetch -format gb |head -50

#esearch -db nucleotide -query "${queryiesss}" |  efetch -format gb | wc -l
esearch -db nucleotide -query "${queryiesss}" |  efetch -format gb -seq_start ${startsss} -seq_stop ${endsss} | wc -l

##-------------------------------------
##do entrez
##-------------------------------------

queryiesss=1015427617
startsss=10
endsss=2000
rm -r /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/
mkdir -p /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/
for clustersss in $(cut -f 1 /data/projects/p539_crisprscope/CRISPRblast//blast_02/non_crispr_or_plasmid_blastHITS_curated.txt)
do
#echo ${clustersss}

queryiesss=$(grep "^${clustersss}$(printf '\t')"  /data/projects/p539_crisprscope/CRISPRblast//blast_02/non_crispr_or_plasmid_blastHITS_curated.txt |cut -f 2|cut -d '|' -f 2 )
startsss=$(grep "^${clustersss}$(printf '\t')"  /data/projects/p539_crisprscope/CRISPRblast//blast_02/non_crispr_or_plasmid_blastHITS_curated.txt |cut -f 9)
endsss=$(grep "^${clustersss}$(printf '\t')"  /data/projects/p539_crisprscope/CRISPRblast//blast_02/non_crispr_or_plasmid_blastHITS_curated.txt |cut -f 10)




 if [ "$startsss" -gt "$endsss"  ]
        then
        #echo -e "reversed"
        startsss_cor=$(echo $endsss|awk '{print $1-2000}')
        endsss_cor=$(echo $startsss|awk '{print $1+2000}')
        
      else
        # echo -e "correct"
        startsss_cor=$(echo $startsss|awk '{print $1-2000}')
        endsss_cor=$(echo $endsss|awk '{print $1+2000}')
      fi
      
      
   if [ "$startsss_cor" -lt "0"  ]
        then
        startsss_cor=1
      fi

echo -e ${clustersss}"\t"${queryiesss}"\t"${startsss_cor}"\t"${endsss_cor}

esearch -db nucleotide -query "${queryiesss}" |  efetch -format gb -seq_start ${startsss_cor} -seq_stop ${endsss} > /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/${clustersss}_${queryiesss}.txt


done


##--------------------------
### which are CRISPR
##--------------------------
ls /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/|wc -l 
rm /data/projects/p539_crisprscope/CRISPRblast//blast_02/tmp.txt
for filesss in $(ls /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/)
do
grep "FEATURES" -A 1000  /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/$filesss |grep "CRISPR" -i |head -1 >> /data/projects/p539_crisprscope/CRISPRblast//blast_02/tmp.txt
done
wc -l /data/projects/p539_crisprscope/CRISPRblast//blast_02/tmp.txt

##--------------------------
###  make nice output
##--------------------------
rm /data/projects/p539_crisprscope/CRISPRblast//blast_02/CRISPRblast_nt_vincent_CRISPR.txt
for filesss in $(ls /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/)
do
countss=$(grep "FEATURES" -A 1000  /data/projects/p539_crisprscope/CRISPRblast//blast_02/efetch/$filesss |grep "CRISPR" -i | wc -l )

clustersss=$(echo ${filesss}|cut -d '_' -f 1)

echo -e ${clustersss}"\t"${countss}
   if [ "$countss" -gt 0  ]
        then
        
        echo -e ${clustersss}"\t"${countss}"\tCRISPRCRISPRCRISPR"
        
        grep "^${clustersss}$(printf '\t')" /data/projects/p539_crisprscope/CRISPRblast//blast_02/non_crispr_or_plasmid_blastHITS_curated.txt >> /data/projects/p539_crisprscope/CRISPRblast//blast_02/CRISPRblast_nt_vincent_CRISPR.txt
        
        
      fi

done

```

# Supplment

### Supplementary infos CRISPR

```{r }
###------------------------------------------------
#number of arrays
###------------------------------------------------

raw_org_arr_sp_class_type <- read_csv("data_submit/raw_org_arr_sp_class_type.csv") %>% group_by(Organism,ArrayID,Cas_class,Cas_subtype) %>% summarise(spacerCounts=n())
raw_org_arr_sp_class_type$strain <- raw_org_arr_sp_class_type$ArrayID %>% gsub("_3$","",.)%>% gsub("_2$","",.)%>% gsub("_1$","",.)%>% gsub("_4$","",.)%>% gsub("_4$","",.)%>% gsub("_5$","",.)%>% gsub("_6$","",.)%>% gsub("_7$","",.)

arry_counts_strain <- raw_org_arr_sp_class_type %>% group_by(Organism,strain) %>% summarise(arrays_num=n())
table(arry_counts_strain$arrays_num)
table(arry_counts_strain$Organism)

arry_counts_strain %>% filter(Organism=="Streptococcus_thermophilus")

arry_counts_strain_sum <- arry_counts_strain %>% group_by(Organism) %>% summarise(mean=mean(arrays_num),median=median(arrays_num),sd=sd(arrays_num))

ggplot(arry_counts_strain,aes(x=Organism,y=arrays_num))+geom_boxplot()+lims(y=c(0,5))


ggplot(arry_counts_strain_sum) +
  geom_bar( aes(x=Organism, y=mean), stat="identity", fill="forestgreen", alpha=0.5) +
  geom_errorbar( aes(x=Organism, ymin=mean-sd, ymax=mean+sd), alpha=0.5) +
  # geom_errorbar( aes(x=Organism, ymin=mean-sd, ymax=mean+sd), width=0.4, colour="orange", alpha=0.9, size=1.5) +
  ggtitle("using standard deviation")+theme(axis.text.y = element_text(size = 12),
      axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1),legend.title = element_blank())

##---------new

NbArrayPerStrain <- read_csv("data_submit/NbArrayPerStrain.csv")

table(NbArrayPerStrain$Organism)

NbArrayPerStrain %>% filter(Organism=="Streptococcus_thermophilus")

ggplot(NbArrayPerStrain,aes(x=Organism,y=nb_array_strain))+geom_boxplot()+theme_classic()+labs(x="",y="count")+theme(axis.text.y = element_text(size = 12),
      axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1),legend.title = element_blank())


NbArrayPerStrain_means <- NbArrayPerStrain %>% group_by(Organism) %>% summarise(mean=mean(nb_array_strain),median=median(nb_array_strain),sd=sd(nb_array_strain))

ggplot(NbArrayPerStrain_means) +
  geom_bar( aes(x=Organism, y=mean), stat="identity", fill="forestgreen", alpha=0.5) +
  geom_errorbar( aes(x=Organism, ymin=mean-sd, ymax=mean+sd), alpha=0.5) +
  # geom_errorbar( aes(x=Organism, ymin=mean-sd, ymax=mean+sd), width=0.4, colour="orange", alpha=0.9, size=1.5) +
  ggtitle("using standard deviation")+theme(axis.text.y = element_text(size = 12),
      axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1),legend.title = element_blank())

###-------add to phylogeny

NbArrayPerStrain_means_cor <- NbArrayPerStrain_means
NbArrayPerStrain_means_cor$id <- gsub("_"," ",NbArrayPerStrain_means_cor$Organism)
NbArrayPerStrain_means_cor <- NbArrayPerStrain_means_cor %>% select(-Organism) %>% rename(value=mean) %>% add_column(category="all") %>% select(id,value,category,sd)

table(pathway_identification_species_plot_complete$species)
table(NbArrayPerStrain_means$species)
NbArrayPerStrain_means$species %in% pathway_identification_species_plot_complete$species



p8 <- facet_plot(gtree_01, panel = 'defense systems', data = NbArrayPerStrain_means,
                geom = geom_tile,color = "grey",width=0.7,height=0.7,size=0.2,
                mapping = aes(x = mechanism_final,fill=percent_species))+scale_x_discrete()+scale_fill_gradient2(na.value="white",low = "white", high = "red", mid = "white",
   midpoint = 0, limit = c(-100,100), space = "Lab",
   name="max growth")+theme(axis.text.y = element_text(size=7),
      axis.text.x = element_text(angle = 45,size=8,hjust=1),axis.title=element_blank(),
    legend.title = element_blank()) #+xlim_expand(c(0,500), 'Genome count')
p8


p9 <- facet_plot(gtree_01, panel = 'Genome count', data = NbArrayPerStrain_means_cor, 
                geom = geom_barh,fill="grey66",
                mapping = aes(x = value), width = 0.7, 
                stat='identity')+ xlim_expand(xlim=c(0, 6.7), panel = "Genome count") 
                # geom_errorbarh aes(xmin=value-sd, xmax=value+sd), alpha=0.5)
p9


max(NbArrayPerStrain_means_cor$value+NbArrayPerStrain_means_cor$sd)

p10 <- facet_plot(p9, panel = 'Genome count_02', data = NbArrayPerStrain_means_cor, 
                geom = geom_errorbarh, 
                mapping = aes(x=value,xmin=value-sd, xmax=value+sd), width = 0.7)+ xlim_expand(xlim=c(0, 6.7), panel = "Genome count_02") 
                # geom_errorbarh aes(xmin=value-sd, xmax=value+sd), alpha=0.5)
p10

###---------------------------------
##number of CRISPR spacers
###---------------------------------

raw_org_arr_sp_class_type <- read_csv("data_submit/raw_org_arr_sp_class_type.csv") %>% group_by(Organism,ArrayID,Cas_class,Cas_subtype) %>% summarise(spacerCounts=n()) %>% mutate(Organism=gsub("_"," ",Organism)) %>% filter(spacerCounts<100)

ggplot(raw_org_arr_sp_class_type,aes(x=spacerCounts)) +
  geom_bar() +
  xlab("Number of spacers") +
  ylab("Array count")+
  theme_classic()

ggplot(raw_org_arr_sp_class_type,aes(x=Organism,y=spacerCounts)) +
  geom_boxplot() +
  xlab("") +
  ylab("Array count")+
  theme_classic()+coord_trans(y="log10")

raw_org_arr_sp_class_type_cleaned <- raw_org_arr_sp_class_type %>% rename(id=Organism)%>% rename(value=spacerCounts) %>% add_column(category="all") %>% ungroup() %>% select(id,value,category)


table(pathway_identification_species_plot_complete$species)
table(raw_org_arr_sp_class_type_cleaned$id)
raw_org_arr_sp_class_type_cleaned$id %in% pathway_identification_species_plot_complete$species
sum(!raw_org_arr_sp_class_type_cleaned$id %in% gtree_01$data$label)

p11 <- facet_plot(p10, panel = 'array size', data = raw_org_arr_sp_class_type_cleaned, 
                geom = geom_boxploth, 
                mapping = aes(x= log(value),group=label),outlier.shape=NA)+ xlim_expand(xlim=c(log(), 4.60517), panel = "array size")#+scale_x_range(breaks=c(2.3,3.2,3.9,4.6), panel = "array size")
                # geom_errorbarh aes(xmin=value-sd, xmax=value+sd), alpha=0.5)
p11


mean(raw_org_arr_sp_class_type_cleaned$value)
mean(NbArrayPerStrain_means_cor$value)

```


### Assign to environment

```{r}
##-------------------------------------
#assign samples to mesophilic or thermophilic
##-------------------------------------

data_environment_species <- OTU_subset_long_extended_full_dominance_again_final_02 %>% filter(species_new %in% c("Streptococcus thermophilus","Lactococcus lactis")) %>% select(sample_name,species_new,relAbundance_02) %>% spread(., species_new, relAbundance_02)%>%replace(is.na(.), 0) %>% mutate(total_dom=`Lactococcus lactis`+`Streptococcus thermophilus`) %>% filter(total_dom>5)
# data_environment_species$`Streptococcus thermophilus`
ggplot(data_environment_species,aes(x=total_dom))+geom_histogram()+theme_classic()


metagenome_repeats_by_projects_wide <- spread(metagenome_repeats_by_projects_long, cluster_repeat_identity, proxy)%>%replace(is.na(.), 0)  %>% remove_rownames()%>% column_to_rownames(., var = "sample_name")# %>% remove_rownames %>% column_to_rownames(var="spacer_name") 

data_environment_species$environment <- ifelse(data_environment_species$`Lactococcus lactis`>data_environment_species$`Streptococcus thermophilus`,"mesophilic","thermophilic")

table(data_environment_species$environment)

data_environment_species$diff <- abs(data_environment_species$`Lactococcus lactis`-data_environment_species$`Streptococcus thermophilus`)-data_environment_species$total_dom

ggplot(data_environment_species,aes(x=diff))+geom_histogram()+theme_classic()


data_environment_species_environ <- data_environment_species %>% select(sample_name,environment)

##-------------------------------------
#see if the species are mor abundant in either or
##-------------------------------------

data_environment_allspecies_env <- OTU_subset_long_extended_full_dominance_again_final_02 %>% select(sample_name,species_new,relAbundance_02) %>% inner_join(.,data_environment_species_environ,by="sample_name")

ggplot(data_environment_allspecies_env,aes(x=environment,y=relAbundance_02))+geom_boxplot()+theme_classic()+facet_wrap(~species_new,scales = "free_y")

table(data_environment_allspecies_env$species_new) %>% as.data.frame() %>% filter(Freq<8)
table(data_environment_allspecies_env$species_new) %>% as.data.frame() %>% filter(Freq>7)


ggplot(data_environment_allspecies_env,aes(x=environment,y=relAbundance_02))+geom_boxplot(alpha=0.2)+theme_classic()+facet_wrap(~species_new,scales = "free_y")+geom_signif( comparisons = list(c("mesophilic", "thermophilic")),map_signif_level = FALSE )+geom_jitter(alpha=0.8)

# plot_phage_cluster <- ggplot(IMGVR_Hits_join_reduced_counts_depupli_sterm_sub_FINAL,aes(x=EXPLAINEd,y=count))+geom_boxplot()+theme_classic()+ stat_compare_means()+labs(y="Phage vOTU size",x="")

#+geom_signif( comparisons = list(c("unobserved", "CRISPR match")),map_signif_level = FALSE )
plot_phage_cluster

###-------------------
#thermophilic samples
c("Lactobacillus zeae","Lactobacillus delbrueckii","Streptococcus thermophilus","Lactobacillus helveticus")
unique(data_environment_allspecies_env$species_new)
#mesophilic samples
c("Lactobacillus versmoldensis","Lactococcus lactis","Streptococcus thermophilus","Propionibacterium freudenreichii")

##------------------------
###color plots
##------------------------
data_environment_allspecies_env$species_assignement <- ifelse(data_environment_allspecies_env$species_new%in%c("Lactobacillus zeae","Lactobacillus delbrueckii","Streptococcus thermophilus","Lactobacillus helveticus"),"thermophilic",ifelse(data_environment_allspecies_env$species_new%in%c("Lactobacillus versmoldensis","Lactococcus lactis","Streptococcus thermophilus","Propionibacterium freudenreichii"),"mesophilic","generalist"))

data_environment_allspecies_env$species_assignement = factor(data_environment_allspecies_env$species_assignement, levels=rev(c("generalist","mesophilic","thermophilic")))

colorssss <- c("orange","blue","grey77")

ggplot(data_environment_allspecies_env,aes(x=environment,y=relAbundance_02,fill=species_assignement,color=species_assignement))+geom_boxplot(alpha=0.2)+theme_classic()+facet_wrap(~species_new,scales = "free_y")+geom_signif( comparisons = list(c("mesophilic", "thermophilic")),map_signif_level = FALSE )+geom_jitter(alpha=0.8)+scale_fill_manual(values = colorssss)+scale_color_manual(values = colorssss)

data_environment_allspecies_env$species_assignement <- ifelse(data_environment_allspecies_env$species_new%in%c("Lacticaseibacillus casei","Lactobacillus delbrueckii","Streptococcus thermophilus","Lactobacillus helveticus"),"thermophilic",ifelse(data_environment_allspecies_env$species_new%in%c("Companilactobacillus versmoldensis","Lactococcus lactis","Streptococcus thermophilus","Propionibacterium freudenreichii"),"mesophilic","generalist"))


###--------------------------------------------
#change names

data_environment_allspecies_env$species_new <- plyr::revalue(data_environment_allspecies_env$species_new,c("Lactobacillus rhamnosus"        = "Lacticaseibacillus rhamnosus",
"Lactobacillus versmoldensis"    = "Companilactobacillus versmoldensis",
"Lactobacillus curvatus"         = "Latilactobacillus curvatus",
"Lactobacillus rhamnosus"        = "Lacticaseibacillus rhamnosus",
"Lactobacillus coryniformis"     = "Loigolactobacillus coryniformis",
"Lactobacillus fermentum"        = "Limosilactobacillus fermentum",
"Lactobacillus zeae"             = "Lacticaseibacillus casei"))


final_abudnance_plot <- ggplot(data_environment_allspecies_env,aes(x=environment,y=relAbundance_02,fill=species_assignement,color=species_assignement))+geom_boxplot(alpha=0.2)+theme_classic()+facet_wrap(~species_new,scales = "free_y")+geom_signif( comparisons = list(c("mesophilic", "thermophilic")),map_signif_level = FALSE )+geom_jitter(alpha=0.8)+scale_fill_manual(values = colorssss)+scale_color_manual(values = colorssss)+labs(x="Metagenomic sample environement",y="Relative abundance")

final_abudnance_plot
  png("~/Desktop/Projects/2020_CRISPRscope/Plots/species_environement_assignement.png", width = 3500, height = 2100,res=300)
   # pdf("~/Desktop/Projects/2020_CRISPRscope/Plots//metagenome_only_repeats_subtype.pdf",width=6,height=3)

final_abudnance_plot

dev.off()

```

## ANTI-CRISPR

Get ANTI-CRISPRs collection. 
```{bash}
cd /home/vincent/Desktop/Projects/2020_CRISPRscope/Anti-CRISPR

sed '1d' anti-CRISPR_assembly.csv |awk -F "\t" '{OFS="\n"}{print ">"$1"_"$2,$6}'| sed 's/ //g' > anti-CRISPR_assembly.fasta

scp -r /home/vincent/Desktop/Projects/2020_CRISPRscope/Anti-CRISPR/anti-CRISPR_assembly.fasta vincent@130.223.51.66:/data/projects/p539_crisprscope/CRISPRblast/
scp -r /home/vincent/Downloads/IMGVR_all_proteins.faa.gz vincent@130.223.51.66:/data/projects/p539_crisprscope/CRISPRblast/

cd /data/projects/p539_crisprscope/CRISPRblast/
gunzip -c IMGVR_all_proteins.faa.gz > /work/Databases/20210128_img_vr/IMGVR_all_proteins.faa
```

cd-hit with anti-crispr collection

```{bash}
sed 's/>/>ANTICRISPR_/g' /data/projects/p539_crisprscope/CRISPRblast/anti-CRISPR_assembly.fasta >> /work/Databases/20210128_img_vr/IMGVR_all_proteins.faa
 cd /data/projects/p539_crisprscope/CRISPRblast/blast_antiCRISPR/
 
 mmseqs easy-cluster /work/Databases/20210128_img_vr/IMGVR_all_proteins.faa /work/Databases/20210128_img_vr/results /work/Databases/20210128_img_vr/tmp

##-------------------cluster
less /work/Databases/20210128_img_vr/results_cluster.tsv

grep "ANTICRISPR_" -c /work/Databases/20210128_img_vr/results_cluster.tsv
grep "ANTICRISPR_"  /work/Databases/20210128_img_vr/results_cluster.tsv |cut -f 1|sort|uniq |wc -l

rm /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs.tsv
for clustersss in $(grep "ANTICRISPR_"  /work/Databases/20210128_img_vr/results_cluster.tsv |cut -f 1|sort|uniq)
do
echo ${clustersss}
grep "${clustersss}" /work/Databases/20210128_img_vr/results_cluster.tsv >> /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs.tsv

done

##---------------------------------

grep "IMGVR_UViG_2537561701_000001|2537583118|2538309733"  /work/Databases/20210128_img_vr/IMGVR_all_Sequence_information.tsv

grep "IMGVR_UViG_2537561701_000001"  /work/Databases/20210128_img_vr/IMGVR_all_Sequence_information.tsv
grep "UGV-GENOME-0217557"  /work/Databases/20210128_img_vr/IMGVR_all_Sequence_information.tsv

cut -f 2 /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs.tsv |grep "ANTICRISPR" -v |cut -d '|' -f 1 >  /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs_REFs.tsv

 grep "UGV-GENOME-0217557"  /work/Databases/20210128_img_vr/IMGVR_all_Sequence_information.tsv|cut -f 6
 
 ##--------identify vOTUs with antiCRISPRs
 rm /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs_vOTUs.tsv
for refsss in $(cat /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs_REFs.tsv)
do

 awk -F "\t" -v refzzz="$refsss" '{if($1==refzzz)print $6}'  /work/Databases/20210128_img_vr/IMGVR_all_Sequence_information.tsv >> /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs_vOTUs.tsv


done

wc -l /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs_REFs.tsv
wc -l /work/Databases/20210128_img_vr/results_cluster_only_ANTICRISPRs_vOTUs.tsv

```


```{r}
###---------------------
 IMGVR_Hits_join <- read_csv("data_submit/IMGVR_Hits_join.csv")

IMGVR_Hits_join_reduced <- IMGVR_Hits_join %>% select(c("qseqid", "sseqid" ,"vOTU","Prediction method","Host taxonomy prediction" )) %>% filter(qseqid %in% IMGVR_Hits_join_short_filter_sipho$qseqid)
IMGVR_Hits_join_reduced$qseqid %>% unique() %>% length()

IMGVR_Hits_join_reduced$taxmap1 <- str_split_fixed(IMGVR_Hits_join_reduced$`Host taxonomy prediction`, fixed(";"), 10)[,7] 

IMGVR_Hits_join_reduced_counts <- IMGVR_Hits_join_reduced %>% group_by(qseqid,sseqid,vOTU,taxmap1) %>% summarise(counts=n()) %>% arrange(., qseqid, -counts)

IMGVR_Hits_join_reduced_counts_depupli <- IMGVR_Hits_join_reduced_counts[!duplicated(IMGVR_Hits_join_reduced_counts[,c(-4,-5)]), ]

table(IMGVR_Hits_join_reduced_counts_depupli$taxmap1)

vOTUs_with_StermHits <- IMGVR_Hits_join_reduced_counts_depupli %>% filter(taxmap1=="Streptococcus thermophilus") %>% select(vOTU) %>% ungroup() %>% distinct(vOTU)

###-----------------all Sterm vOTUs

vOTUS_in_which_species <- read_csv("data_submit/vOTUS_in_which_species.csv")
vOTUs_sterm <- vOTUS_in_which_species %>% filter(taxon=="Streptococcus thermophilus") %>% select(vOTU) %>% distinct()
###-------bring together

vOTUs_with_StermHits$vOTU %in% vOTUs_sterm$vOTU

vOTUs_sterm$group <- ifelse(vOTUs_sterm$vOTU%in%vOTUs_with_StermHits$vOTU,"CRISPR targeted","no hit")
##---
vOTUS_with_ANTICRISPR <- read_csv("data_submit/results_cluster_only_ANTICRISPRs_vOTUs.tsv", col_names = "vOTU")

vOTUs_sterm$ANTICRISPR <- ifelse(vOTUs_sterm$vOTU%in%vOTUS_with_ANTICRISPR$vOTU,"yes","no")

vOTUs_sterm_sum <- vOTUs_sterm %>%  group_by(group,ANTICRISPR) %>% summarise(counts=n())
colnames(vOTUs_sterm_sum)[2] <- "Anti-CRISPR contained"

plot_antiCRISPR <- ggplot(vOTUs_sterm_sum,aes(x=group,y=counts,fill=`Anti-CRISPR contained`))+geom_bar(stat = "identity")+theme_classic()+labs(x="",y="number vOTUs")
plot_antiCRISPR

```


```{bash}
mappingFile=/data/projects/p539_crisprscope/CRISPRblast/anti-CRISPR_assembly.fasta
BaseLocation=/data/projects/p539_crisprscope/CRISPRblast/
mkdir -p ${BaseLocation}/blast_antiCRISPR

cd /home/vincent/anaconda3/envs/PROKKA/db/hmm
#makeblastdb -dbtype nucl -in IMGVR_all_nucleotides.fna

tblastn -num_threads 37 -max_hsps 1 -max_target_seqs 1 -show_gis \
  -query ${mappingFile} \
  -outfmt "6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle slen" \
  -db /work/Databases/20210128_img_vr/IMGVR_all_nucleotides.fna \
  -out  ${BaseLocation}/blast_antiCRISPR/ANTI_CRISPRblast_IMGVR.txt \
  -evalue 0.01 -word_size 4

wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB.txt 
wc -l   ${BaseLocation}/blast/CRISPRblast_ViralDB_old.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_nucleotideDB.txt
wc -l   ${BaseLocation}/blast/CRISPRblast_IMGVR.txt

```



## cOTU vs. spacer correlation

```{r}
export_fig4A_values <- read_csv("data_submit/export_fig4A_values.csv") %>% filter(!name %in% c( "Lordan et al. 2019  ( 6 )",       "Pasolli et al. 2020  ( 1594 )" ,  "Swiss raclette  ( 2427 )" ,       "Swiss starter culture  ( 2882 )", "Walsh et al. 2020  ( 1607 )"    )) %>% filter(final_hit=="Viral") %>%  mutate(non_viral=100-HitRatio) %>% select(name,non_viral)

fraction_votu_per_species <- read_csv("data_submit/fraction_votu_per_species.csv") %>% select(Organism,n_votus_per_total)

export_fig4A_values$name <- plyr::revalue(export_fig4A_values$name,c("Bani  ( 606 )" ="Bifidobacterium animalis","Cacn  ( 38 )"="Cutibacterium acnes","Ccas  ( 140 )"="Corynebacterium casei","Laci  ( 59 )"="Lactobacillus acidophilus","Lcas  ( 301 )"="Lacticaseibacillus casei","Lcor  ( 400 )"="Loigolactobacillus coryniformis","Lcur  ( 545 )"="Latilactobacillus curvatus","Ldel  ( 3189 )"="Lactobacillus delbrueckii","Leulac  ( 163 )"="Leuconostoc lactis","Lfer  ( 1991 )"="Limosilactobacillus fermentum","Lhelv  ( 1590 )"="Lactobacillus helveticus","Lpla  ( 567 )"="Lactiplantibacillus plantarum","Lpseu  ( 15 )"="Leuconostoc pseudomesenteroides","Lrham  ( 372 )"="Lacticaseibacillus rhamnosus","Msciu  ( 118 )"="Mammaliicoccus sciuri","Pfreu  ( 2309 )"="Propionibacterium freudenreichii","Ppar  ( 258 )"="Pediococcus parvulus","Ppen  ( 137 )"="Pediococcus pentosaceus","Sequo  ( 16 )"="Staphylococcus equorum","Slute  ( 316 )"="Streptococcus lutetiensis","Stherm  ( 3376 )"="Streptococcus thermophilus"))

final_explaine_votu <- merge(export_fig4A_values,fraction_votu_per_species,by.x="name",by.y="Organism")

ggplot(final_explaine_votu,aes(x=non_viral,y=n_votus_per_total))+geom_point()+theme_classic()

```

## HMM ABI

```{r}

all_scans_ABI <- read_delim("data_submit/all_scans_ABI.txt", "\t", escape_double = FALSE, col_names = c("genome","geneOfInterest","gene")) %>% filter(!geneOfInterest %in% c("GalA","GalB","AbiL","AbiK","AbiA")) %>% filter(!geneOfInterest %in% c("STM_4041","STM14_4038","STM14_4040"))
all_scans_ABI$geneOfInterest <- plyr::revalue(all_scans_ABI$geneOfInterest,c("JatA"="JetA"))

all_scans_ABI_count <- all_scans_ABI %>% group_by(genome,geneOfInterest) %>% dplyr::summarise(counts=n()) %>% mutate(mechanism_final=geneOfInterest)%>% mutate(mechanism_gene_count=1)

##-----------------------------------

see_gene_count_ABI <- all_scans_ABI_count 

tmp <- see_gene_count_ABI %>% filter(is.na(mechanism_final))

pathway_identification_ABI <- see_gene_count_ABI %>% group_by(genome,mechanism_final,mechanism_gene_count) %>%  summarise(number_genes_identified=n())

pathway_identification_ABI$percentPathway <- 100*(pathway_identification_ABI$number_genes_identified/pathway_identification_ABI$mechanism_gene_count)

pathway_identification_ABI$Completness <- ifelse(pathway_identification_ABI$percentPathway>99,"Yes","No")

pathway_identification_ABI$Genus <- str_split_fixed(pathway_identification_ABI$genome, fixed("_"), 3)[,1]
pathway_identification_ABI$spec <- str_split_fixed(pathway_identification_ABI$genome, fixed("_"), 3)[,2]
pathway_identification_ABI$species <- paste(pathway_identification_ABI$Genus,pathway_identification_ABI$spec,sep="_")

###------------------------------genomes counts
species_counts_ABI <- all_scans_ABI %>% select(genome) %>% mutate(species=paste(str_split_fixed(genome, fixed("_"), 3)[,1],str_split_fixed(genome, fixed("_"), 3)[,2],sep="_")) %>% unique() %>% group_by(species) %>% summarise(counts=n())

###------------------------------abundances
tmp1 <- pathway_identification_ABI %>% group_by(species,mechanism_final) %>% summarise(Present=sum(Completness=="Yes"),Absent=sum(Completness!="Yes"))

pathway_identification_species_ABI <- merge(tmp1,species_counts_ABI) %>% mutate(percent_species=100*(Present/counts))

##-------------------------------

ggheatmap <- ggplot(pathway_identification_species_ABI, aes(y=species, x=mechanism_final, fill = percent_species))+
 geom_tile(color = "white",size=1.1)+
 scale_fill_gradient2(low = "white", high = "red", mid = "white",
   midpoint = 0, limit = c(-100,100), space = "Lab",
   name="max growth") +theme_classic()+
  scale_x_discrete(position = "top") +
 theme(axis.text.y = element_text(size=7),
      axis.text.x = element_text(angle = 45,vjust=1,
    size = 12, hjust = 0),axis.title=element_blank(), 
    legend.title = element_blank())

 print(ggheatmap)
 
 ##-------------------------------
 
 
 tmps_02 <- pathway_identification_ABI %>% group_by(species,genome) %>% summarise(number_mechanism=sum(Completness=="Yes"))

 plot_ABI <- ggplot(tmps_02,aes(x=species,y=number_mechanism))+geom_boxplot()+theme_classic()+theme(axis.text.x = element_text(angle = 75,size=8,hjust=1))+lims(y=c(0,6))+labs(x="",y="number ABI")
 
 plot_ABI
 
  ##-------------------------------heatmap all ABI

  ##find number of strains
  tmptmp <- pathway_identification %>% ungroup()%>% select(genome,species) %>% distinct()
tmp_04 <- table(tmptmp$species) %>% as.data.frame() %>% rename(., "tot_strains"="Freq") %>% rename(., "species"="Var1")
  

  tmps_03 <- pathway_identification_ABI %>% group_by(species,genome) %>% summarise(number_mechanism=sum(Completness=="Yes"),Absent=sum(Completness!="Yes")) %>% group_by(species) %>% summarise(genomes_with_ABI=n()) %>% inner_join(.,tmp_04,by="species") %>% mutate(percent_ABIs=100*(genomes_with_ABI/tot_strains))

   tmps_02tmp1 <- pathway_identification_ABI %>% group_by(species,mechanism_final) %>% summarise(Present=sum(Completness=="Yes"),Absent=sum(Completness!="Yes"))
```

### Plasmid representation

Here, we check if plasmid and bacterial hits are over represented in the protospacer hits. 

```{bash}

sed 's/ ,/,/g' ~/Desktop/Projects/2020_CRISPRscope/data/meta_sp_protosp_coverage_hits.csv > ~/Desktop/Projects/2020_CRISPRscope/data/meta_sp_protosp_coverage_hits_corrected.csv

```


```{r}
meta_sp_protosp_coverage_hits <- read_csv("data_submit/meta_sp_protosp_coverage_hits_corrected.csv")

meta_sp_protosp_coverage_hits <- read.csv("data_submit/meta_sp_protosp_coverage_hits.csv")  

sum(meta_sp_protosp_coverage_hits$protospacer_cov<1)
unique(meta_sp_protosp_coverage_hits$sample)
unique(meta_sp_protosp_coverage_hits$cluster_spacer_identity) %>% length()

###-----------------------------
#bacteria
databaseINterest="nt hit"
databaseINterest="Plasmid"
databaseINterest="Viral"
databaseINterest="No hit"

# table(meta_sp_protosp_coverage_hits$final_hit)

# Interest_nometa <- meta_sp_protosp_coverage_hits %>% filter(final_hit==databaseINterest) %>% filter(protospacer_cov>0) %>% select(cluster_spacer_identity) %>% distinct()

Interest_meeta <- meta_sp_protosp_coverage_hits %>% filter(final_hit==databaseINterest) %>% filter(protospacer_cov>0) %>% select(sample,cluster_spacer_identity) %>% unique() %>% nrow()

Interest_nometa <- meta_sp_protosp_coverage_hits %>% filter(final_hit==databaseINterest) %>% filter(protospacer_cov==0) %>% select(sample,cluster_spacer_identity) %>% unique() %>% nrow()

noInterest_meeta <- meta_sp_protosp_coverage_hits %>% filter(final_hit!=databaseINterest) %>% filter(protospacer_cov>0) %>% select(sample,cluster_spacer_identity) %>% unique() %>% nrow()

noInterest_nometa <- meta_sp_protosp_coverage_hits %>% filter(final_hit!=databaseINterest) %>% filter(protospacer_cov==0) %>% select(sample,cluster_spacer_identity) %>% unique() %>% nrow()

fisher_prep <- data.frame(present=c(Interest_meeta,noInterest_meeta),absent=c(Interest_nometa,noInterest_nometa))

rownames(fisher_prep) <- c(databaseINterest,"rest")
fisher_prep %>% mutate(percent=100*(present/(present+absent)))
fisher_prep 
test <- fisher.test(fisher_prep)
test
```


